---
title: "Hands-on Exercise 6"
description: |
  In this hands-on exercise, we learn how to derive spatial weights using functions provided by the spdep package. We also learn how to apply these spatial weights to compute geographically referenced atributes.
author:
  - name: Genice Goh
    url: {}
date: 09-18-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Importing and Loading Packages

```{r echo=TRUE}
packages <- c('spdep', 'sf', 'tmap', 'tidyverse')
for (p in packages) {
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Importing Data

*st_read()* of **sf** package is used to import the geospatial data, which is in Shapefile format.

```{r echo=TRUE}
hunan <- st_read(dsn = "data/shapefile", 
                 layer = "Hunan")
```

Since the aspatial data is in CSV format, we use *read_csv()* of **readr** package to import it.

```{r echo=TRUE}
hunan2012 <- read_csv("data/attribute/Hunan_2012.csv")
```

## Performing Relational Join

*left_join()* of **dplyr** package is used to update the attribute table of `hunan`’s SpatialPolygonsDataFrame with the attribute fields of `hunan2012` data frame.

```{r echo=TRUE}
hunan <- left_join(hunan,hunan2012)
```

## Visualising Regional Development Indicator

We are going to prepare a choropleth map showing the distribution of GDPPC 2012 by using *qtm()* of **tmap** package.

```{r echo=TRUE, fig.width=10, fig.height=5}
qtm(hunan, "GDPPC")
```

## Computing Contiguity Spatial Weights

We will learn how to use *poly2nb()* of **spdep** package to compute contiguity weight matrices for the study area. This function builds a neighbours list *based on regions with contiguous boundaries (i.e. common borders)*. 

We can pass a “queen” argument that takes values **TRUE (default)** or **FALSE**. If we don’t specify queen = FALSE, this function will return a list of *first order neighbours* using the **Queen** criteria.

### Computing (QUEEN) contiguity based neighbours

```{r echo=TRUE}
wm_q <- poly2nb(hunan, queen=TRUE)
summary(wm_q)
```

The summary report above shows that there are *88 area units* in Hunan. The most connected area unit has *11 neighbours*. There are *two area units with only one heighbours*.

For each polygon in the polygon object, wm_q lists all neighboring polygons. To see the neighbors for the first polygon in the object:

```{r echo=TRUE}
wm_q[[1]]
```

Polygon 1 has *5 neighbors*. The numbers represent the *polygon IDs* as stored in `hunan` SpatialPolygonsDataFrame class.

<br>

We can retrieve the county name of Polygon ID = 1 using the code chunk below:

```{r echo=TRUE}
hunan$County[1]
```

The output reveals that Polygon ID=1 is Anxiang county.

<br>

To reveal the county names of the five neighboring polygons, the code chunk will be used:

```{r echo=TRUE}
hunan$NAME_3[c(2,3,4,57,85)]
```

<br>

We can display the complete weight matrix by using *str()*.

```{r echo=TRUE}
str(wm_q)
```

Note: The report can be quite long!! 

### Creating (ROOK) contiguity based neighbours

```{r echo=TRUE}
wm_r <- poly2nb(hunan, queen=FALSE)
summary(wm_r)
```

The summary report above shows that there are *88 area units* in Hunan. The most connected area unit has *10 neighbours*. There are *two area units with only one neighbour*.

## Visualising contiguity weights

A **connectivity** graph *takes a point* and *displays a line to each neighboring point*. Since we are currently only working with polygons, we need to get points in order to make the graphs, and they are most typically polygon centroids. 

### Getting Latitude and Longitude of Polygon Centroids

We will need points to associate with each polygon before we can make our connectivity graph and the coordinates need to be in a separate data frame for this to work. 

To do this, we will use a mapping function which *applies a given function to each element of a vector* and *returns a vector of the same length*. Our input vector will be the geometry column of `hunan` and we will be using *st_centroid()* of **sf** package. We will be using *map_dbl()* variation of *map()* from the **purrr** package and it *returns a double vector*.

To get the longitude values, we map *st_centroid()* over the geometry column of `hunan` and access it using `[[1]]`. This is because the longitude is the first value in each centroid.

```{r echo=TRUE}
longitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[1]])
```

<br>

Similarly, we will retrieve the latitude values by accessing it using `[[2]]`, since the latitude is the second value in each centroid.

```{r echo=TRUE}
latitude <- map_dbl(hunan$geometry, ~st_centroid(.x)[[2]])
```

<br>

*cbind()* is now used to put the longitude and latitude values into the same object.

```{r echo=TRUE}
coords <- cbind(longitude, latitude)
```

<br>

We can now check the first few observations to see if things are formatted correctly. 

```{r echo=TRUE}
head(coords)
```

We now have the coordinates of the polygon centroids in a separate data frame!

### Plotting Queen contiguity based neighbours maps

```{r echo=TRUE, fig.width=8, fig,height=6}
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red")
```

### Plotting Rook contiguity based neighbours maps

```{r echo=TRUE, fig.width=8, fig.height=6}
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

### Comparing Queen and Rook contiguity based neighbour maps

```{r echo=TRUE, fig.width=8, fig.height=6}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(wm_q, coords, pch = 19, cex = 0.6, add = TRUE, col= "red", main="Queen Contiguity")
plot(hunan$geometry, border="lightgrey")
plot(wm_r, coords, pch = 19, cex = 0.6, add = TRUE, col = "red", main="Rook Contiguity")
```

## Computing Distance Based Neighbours

We will learn how to derive distance-based weight matrices using *dnearneigh()* of **spdep** package.

The function identifies neighbours of region points by *Euclidean distance with a distance band defined by the lower d1 and upper d2 bounds*. If unprojected coordinates are used and either (specified in the coordinates object x) or (with x as a two column matrix and longlat=TRUE), great circle distances in **km** will be calculated assuming the *WGS84 reference ellipsoid*.

### Determine the cut-off distance

Firstly, we need to determine the *upper limit* for distance band by using the steps below:

- Return a matrix with the indices of points belonging to the k nearest neighbours of the polygon centroids by using *knearneigh()* of **spdep**.
- Convert the **knn** object into a neighbours list of class **nb** with a list of integer vectors containing neighbour region number ids by using *knn2nb()*.
- Return the length of neighbour relationship edges by using nbdists() of **spdep**. The function returns in the units of the coordinates if they are projected, else in **km**.
- Remove the list structure of the returned object by using *unlist()*.

```{r echo=TRUE}
k1 <- knn2nb(knearneigh(coords))
k1dists <- unlist(nbdists(k1, coords, longlat = TRUE))
summary(k1dists)
```

The summary report shows that the largest first nearest neighbour distance is 61.79 km, so using this as the upper threshold gives certainty that *all units will have at least one neighbour*.

### Computing fixed distance weight matrix

Now, we will compute the distance weight matrix by using *dnearneigh()* as shown in the code chunk below.

```{r echo=TRUE}
wm_d62 <- dnearneigh(coords, 0, 62, longlat = TRUE)
wm_d62
```

<br>

Another way to display the structure of the weight matrix is to combine *table()* and *card()* of **spdep**.

```{r echo=TRUE}
table(hunan$County, card(wm_d62))
```

### Plotting fixed distance weight matrix

Next, we will plot the distance weight matrix by using the code chunk below.

```{r echo=TRUE, fig.width=8, fig.height=6}
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE)
plot(k1, coords, add=TRUE, col="red", length=0.08)
```

The red lines show the links of 1st nearest neighbours and the black lines show the links of neighbours within the cut-off distance of 62km.

Alternatively, we can plot both of them next to each other by using the code chunk below.

```{r echo=TRUE, fig.width=8, fig.height=6}
par(mfrow=c(1,2))
plot(hunan$geometry, border="lightgrey")
plot(k1, coords, add=TRUE, col="red", length=0.08, main="1st nearest neighbours")
plot(hunan$geometry, border="lightgrey")
plot(wm_d62, coords, add=TRUE, pch = 19, cex = 0.6, main="Distance link")
```

### Computing adaptive distance weight matrix

One of the characteristics of fixed distance weight matrix is that *more densely settled areas* (i.e. urban areas) tend to *have more neighbours* and the *less densely settled areas* (i.e. rural counties) tend to *have less neighbours*. Having many neighbours smoothes the neighbour relationship across more neighbours.

It is possible to control the numbers of neighbours directly using k-nearest neighbours, either accepting asymmetric neighbours or imposing symmetry as shown in the code chunk below.

```{r echo=TRUE}
knn6 <- knn2nb(knearneigh(coords, k=6))
knn6
```

### Plotting adaptive distance weight matrix

We can plot the adaptive distance weight matrix using the code chunk below.

```{r echo=TRUE, fig.width=8, fig.height=6}
plot(hunan$geometry, border="lightgrey")
plot(knn6, coords, pch = 19, cex = 0.6, add = TRUE, col = "red")
```

## Weights based on IDW
We will learn how to derive a spatial weight matrix based on the *Inversed Distance method*.

First, we will compute the distances between areas by using *nbdists()* of **spdep**.

```{r echo=TRUE}
dist <- nbdists(wm_q, coords, longlat = TRUE)
ids <- lapply(dist, function(x) 1/(x))
ids
```

## Row-standardised weights matrix

Next, we need to assign weights to each neighboring polygon. In our case, each neighboring polygon will be assigned equal weight (style=“W”). This is accomplished by *assigning the fraction 1/(#ofneighbors) to each neighboring county* then *summing the weighted income values*. 

It has a drawback: polygons along the edges of the study area will *base their lagged values on fewer polygons* thus potentially **over-** or **under-estimating** the *true nature of the spatial autocorrelation* in the data. 

Note: There are other more robust options are available, notably style=“B”.

```{r echo=TRUE}
rswm_q <- nb2listw(wm_q, style="W", zero.policy = TRUE)
rswm_q
```

The **zero.policy=TRUE** option allows for lists of non-neighbors. This should be used with **caution** since the user may not be aware of missing neighbors in their dataset. However, a zero.policy of FALSE would return an error.

To see the weight of the first polygon’s four neighbors:

```{r echo=TRUE}
rswm_q$weights[1]
```

Each neighbor is assigned a 0.2 of the total weight. This means that when R computes the average neighboring income values, each neighbor’s income will be multiplied by 0.2 before being tallied.

Using the same method, we can also derive a row standardised distance weight matrix by using the code chunk below.

```{r echo=TRUE}
rswm_ids <- nb2listw(wm_q, glist=ids, style="B", zero.policy=TRUE)
rswm_ids
```

## Application of Spatial Weight Matrix

We will learn how to create different spatial lagged variables:

- Spatial lag with row-standardized weights
- Spatial window sum

### Spatial lag with row-standardized weights

We will compute the average neighbour GDPPC value for each polygon. These values are often referred to as **spatially lagged values**.

```{r echo=TRUE}
GDPPC.lag <- lag.listw(rswm_q, hunan$GDPPC)
GDPPC.lag
```

We can append the spatially lag GDPPC values onto `hunan` SpatialPolygonDataFrame by using the code chunk below.  

```{r echo=TRUE}
lag.list <- list(hunan$NAME_3, lag.listw(rswm_q, hunan$GDPPC))
lag.res <- as.data.frame(lag.list)
colnames(lag.res) <- c("NAME_3", "lag GDPPC")
hunan <- left_join(hunan,lag.res)
```

Next, we will plot both the GDPPC and spatial lag GDPPC for comparison using the code chunk below.

```{r echo=TRUE, fig.width=10, fig.height=5}
gdppc <- qtm(hunan, "GDPPC")
lag_gdppc <- qtm(hunan, "lag GDPPC")
tmap_arrange(gdppc, lag_gdppc, asp=1, ncol=2)
```

## Spatial window sum

The spatial window sum *uses and includes the diagonal element*. To begin, we will assign knn6 to a new variable because we will directly alter its structure to add the diagonal elements

```{r echo=TRUE}
knn6a <- knn6
```

<br>

To add the *diagonal element* to the neighbour list, we just need to use *include.self()* from **spdep**.

```{r echo=TRUE}
include.self(knn6a)
```

<br>

Next, we will assign binary weights to the neighbour structure that includes the diagonal element.

```{r echo=TRUE}
binary.knn6 <- lapply(knn6a, function(x) 0*x+1)
binary.knn6[1]
```

<br>

Again, we use *nb2listw()* and *glist()* to explicitly assign weight values.

```{r echo=TRUE}
wm_knn6 <- nb2listw(knn6a, glist = binary.knn6, style = "B")
```

<br>

With our new weight structure, we can compute the lag variable with *lag.listw()*.

```{r echo=TRUE}
lag_knn6 <- lag.listw(wm_knn6, hunan$GDPPC)
```

<br>

Next, we will convert the lag variable **listw** object into a dataframe by using *as.data.frame()*.

```{r echo=TRUE}
lag.list.knn6 <- list(hunan$NAME_3, lag.listw(wm_knn6, hunan$GDPPC))
lag_knn6.res <- as.data.frame(lag.list.knn6)
colnames(lag_knn6.res) <- c("NAME_3", "lag_sum GDPPC")
```

Note: The third line renames the field names of **lag_knn6.res** object into *NAME_3* and *lag_sum GDPPC* respectively. 

<br>

Next, the code chunk below will be used to append *lag_sum GDPPC* values onto `hunan` sf dataframe by using *left_join()* of **dplyr** package.  

```{r echo=TRUE}
hunan <- left_join(hunan, lag_knn6.res)
```
<br>

Lastly, *qtm()* of **tmap** package is used to plot the GDPPC and lag_sum GDPPC map next to each other for quick comparison. 

```{r echo=TRUE, fig.width=12, fig.height=10}
gdppc <- qtm(hunan, "GDPPC")
lag_sum_gdppc <- qtm(hunan, "lag_sum GDPPC")
tmap_arrange(gdppc, lag_sum_gdppc, asp=1, ncol=2)
```