---
title: "Hands-on Exercise 4"
description: |
  This hands-on exercise covers the appropriate functions of spatstat package to perform spatial point patterns analysis.
author:
  - name: Genice Goh
    url: {}
date: 09-05-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading Packages

- sf: import, manage and process vector-based geospatial data
- spatstat: perform point pattern analysis 
- raster: read, write, manipulate, analyse and model of gridded spatial data
- maptools: provide tools for manipulating geographic data 
- tmap: plot static point patterns maps or interactive maps

```{r echo=TRUE, eval=TRUE}
packages = c('sf', 'spatstat', 'raster', 'maptools', 'tmap')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Spatial Data Wrangling

### Importing spatial data

*st_read()* of sf package is used to import the geospatial data

```{r echo=TRUE, eval=TRUE}
childcare_sf <- st_read("data/child-care-services-geojson.geojson") %>%
  st_transform(crs=3414)

sg_sf <- st_read(dsn = "data", layer="CostalOutline")

mpsz_sf <- st_read(dsn = "data", 
                   layer = "MP14_SUBZONE_WEB_PL")
```

Important to ensure that *sg_sf* and *mpsz_sf* are projected in the same projection system before using these data for analysis

```{r echo=TRUE, eval=TRUE}
print(st_crs(sg_sf))
```

```{r echo=TRUE, eval=TRUE}
print(st_crs(mpsz_sf))
```

Although both *sg_sf* and *mpsz_sf* are projected in SVY21, the end of both prints indicate that the EPSG is 9001. This is a wrong EPSG code because the correct EPSG code for SVY21 should be 3414. There is therefore a need to assign the correct EPSG codes to *sg_sf* and *mpsz_sf*.

```{r echo=TRUE, eval=TRUE}
sg3414_sf <- st_set_crs(sg_sf, 3414)

mpsz3414_sf <- st_set_crs(mpsz_sf, 3414)
```

### Mapping the geospatial datasets

It is useful to plot a map to show the geospatial data's spatial patterns

```{r echo=TRUE, eval=TRUE}
tm_shape(sg_sf) +
  tm_polygons() +
tm_shape(mpsz_sf) +
  tm_polygons() +
tm_shape(childcare_sf)+
  tm_dots()
```

*Note: All the geospatial layers are within the same map extend. This shows that their referencing system and coordinate values are referring to similar spatial context*

Alternatively, a pin map can be plotted. Its advantages are:

- Navigate and zoom around the map freely
- Query the information of each simple feature by clicking them
- Can change the background of the internet map layer. 3 are currently provided: ESRI.WorldGrayCanvas (default), OpenStreetMap and ESRI.WorldTopoMap


```{r echo=TRUE, eval=TRUE}
tmap_mode('view')
tm_shape(childcare_sf)+
  tm_dots()
```

```{r echo=TRUE, eval=TRUE}
tmap_mode('plot')
```

*Note: Always remember to switch back to plot mode afterwards because each interactive mode will consume a connection. Avoid displaying too many interactive maps in 1 RMarkdown document when publishing on Netlify.*

## Geospatial Data Wrangling

Many geospatial analysis packages require the input geospatial data in sp’s Spatial* classes

### Converting sf data frames to sp’s Spatial* class

*as_Spatial()* of sf package is used to convert the data from sf data frame to sp’s Spatial* class

```{r echo=TRUE, eval=TRUE}
childcare <- as_Spatial(childcare_sf)
mpsz <- as_Spatial(mpsz3414_sf)
sg <- as_Spatial(sg3414_sf)
```

```{r echo=TRUE, eval=TRUE}
childcare
mpsz
sg
```

### Converting the Spatial* class into generic sp format

*spatstat* requires the data to be in **ppp** object form. To convert Spatial* class to **ppp** object, the **Spatial* class** needs to be convert into **Spatial object** first.

```{r echo=TRUE, eval=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
sg_sp <- as(sg, "SpatialPolygons")
```

```{r echo=TRUE, eval=TRUE}
childcare_sp
sg_sp
```

### Converting generic sp format into spatstat’s ppp format
*as.ppp()*of spatstat is used to convert the **Spatial object** into **spatstat’s ppp** object format

```{r echo=TRUE, eval=TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
childcare_ppp
```

<br>

Plotting *childcare_ppp*

```{r echo=TRUE, eval=TRUE}
plot(childcare_ppp)
```

<br>

Summary statistics of *childcare_ppp*

```{r echo=TRUE, eval=TRUE}
summary(childcare_ppp)
```

Note: In spatial point patterns analysis, the presence of duplicates is a significant issue. The statistical methodology used for spatial point patterns processes is based largely on the assumption that the process is simple (i.e. the points cannot be coincident)

### Handling duplicated points

Check for any duplication in *childcare_ppp*

```{r echo=TRUE, eval=TRUE}
any(duplicated(childcare_ppp))
```

*multiplicity()* is used to count number of coincidence points. To know how many locations have >1 point event, refer to code chunk below:

```{r echo=TRUE, eval=TRUE}
sum(multiplicity(childcare_ppp) > 1)
```

The output shows that there are 128 duplicated points. Plotting the data will  allow us to view their locations

```{r echo=TRUE, eval=TRUE}
tmap_mode('view')
tm_shape(childcare) +
  tm_dots(alpha=0.4, 
          size=0.05)
```
```{r echo=TRUE, eval=TRUE}
tmap_mode('plot')
```

To overcome this problem:

- Delete the duplicates (some useful point events will be lost)
- Jittering, which adds a small perturbation to the duplicate points so that they do not occupy the exact same space
- Make each point “unique” and then attach the duplicates of the points to the patterns as marks (as attributes of the points) 
  - Will however need analytical techniques that take into account these marks

Implementing Jittering Approach

```{r echo=TRUE, eval=TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp, 
                             retry=TRUE, 
                             nsim=1, 
                             drop=TRUE)
```

There are no duplicate points now

```{r echo=TRUE, eval=TRUE}
any(duplicated(childcare_ppp_jit))
```

### Creating owin object

Good practice to confine the analysis within a geographical area when doing spatial point patterns analysis. In spatstat, an object called **owin** is specially designed to represent this polygonal region

```{r echo=TRUE, eval=TRUE}
sg_owin <- as(sg_sp, "owin")

plot(sg_owin)
```

### Combining point events object and owin object

Extract childcare centres that are located within Singapore 

```{r echo=TRUE, eval=TRUE}
childcareSG_ppp = childcare_ppp[sg_owin]
```

<br>

The output object combined both the point and polygon feature in 1 **ppp** object class

```{r echo=TRUE, eval=TRUE}
summary(childcareSG_ppp)
```

<br>

Plot the newly derived childcareSG_ppp 

```{r echo=TRUE, eval=TRUE}
plot(childcareSG_ppp)
```

## First-order Spatial Point Patterns Analysis

- Deriving KDE layer for visualising and exploring the intensity of point processes
- Performing Confirmatory Spatial Point Patterns Analysis by using Nearest Neighbour statistics

### Kernel Density Estimation

Kernel density can be computed with *density()* of spatstat using the following configurations:

- bw.diggle() automatic bandwidth selection method. Other recommended methods are bw.CvL(), bw.scott() or bw.ppl()
- Smoothing kernel used is gaussian (default). Other smoothing methods are: “epanechnikov”, “quartic” or “disc”
- The intensity estimate is corrected for edge effect bias. The default is FALSE

```{r echo=TRUE, eval=TRUE}
kde_childcareSG_bw <- density(childcareSG_ppp,
                              sigma=bw.diggle,
                              edge=TRUE,
                            kernel="gaussian") 

plot(kde_childcareSG_bw)
```

The output density values are way too small to comprehend. This is because the default unit of measurement of SVY21 is in m. As a result, the density values computed is in number of points/square metre

<br>

Retrieve bandwidth used to compute KDE layer 

```{r echo=TRUE, eval=TRUE}
bw <- bw.diggle(childcareSG_ppp)
bw
```

### Rescaling KDE values

*rescale()* is used to covert the unit of measurement from m to km

```{r echo=TRUE, eval=TRUE}
childcareSG_ppp.km <- rescale(childcareSG_ppp, 1000, "km")
```

<br>

Re-run *density()* using the rescaled data set and plot the output KDE map

```{r echo=TRUE, eval=TRUE}
kde_childcareSG.bw <- density(childcareSG_ppp.km, 
                              sigma=bw.diggle, 
                              edge=TRUE, 
                              kernel="gaussian")
plot(kde_childcareSG.bw)
```

*Note: Output image looks identical to the earlier version, the only changes are in the data values*

### Working with different automatic bandwidth methods

Besides bw.diggle(), 3 other spatstat functions can be used to determine the bandwidth: bw.CvL(), bw.scott(), and bw.ppl().

Look at the bandwidth return by these automatic bandwidth calculation methods:

```{r echo=TRUE, eval=TRUE}
bw.CvL(childcareSG_ppp.km)
bw.scott(childcareSG_ppp.km)
bw.ppl(childcareSG_ppp.km)
bw.diggle(childcareSG_ppp.km)
```

*bw.ppl()* is suggested because it tends to produce more appropriate values when the pattern consists **predominantly of tight clusters**. However, if the purpose is to **detect a single tight cluster in the midst of random noise**, then *bw.diggle()* method seems to work best

```{r echo=TRUE, eval=TRUE}
kde_childcareSG.ppl <- density(childcareSG_ppp.km, 
                               sigma=bw.ppl, 
                               edge=TRUE,
                               kernel="gaussian")
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "bw.diggle")
plot(kde_childcareSG.ppl, main = "bw.ppl")
```

### 6.1.3 Working with different kernel methods

By default, the kernel method used in density.ppp() is gaussian. There are three other options: Epanechnikov, Quartic and Dics

```{r echo=TRUE, eval=TRUE}
par(mfrow=c(2,2))
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="gaussian"), 
     main="Gaussian")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="epanechnikov"), 
     main="Epanechnikov")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="quartic"), 
     main="Quartic")
plot(density(childcareSG_ppp.km, 
             sigma=bw.ppl, 
             edge=TRUE, 
             kernel="disc"), 
     main="Disc")
```

### Fixed and Adaptive KDE

#### Computing KDE with fixed bandwidth

We will now compute a KDE layer by defining a bandwidth of 600m. Sigma value used is 0.6 because km is the unit of measurement for childcareSG_ppp.km 

```{r echo=TRUE, eval=TRUE}
kde_childcareSG_600 <- density(childcareSG_ppp.km, sigma=0.6, edge=TRUE, kernel="gaussian")
plot(kde_childcareSG_600)
```

#### Computing KDE with adpative bandwidth

Fixed bandwidth method is **very sensitive to highly skew distribution of spatial point patterns over geographical units**. Use adaptive bandwidth to overcome this issue

*density.adaptive()* of spatstat is used to derive adaptive kernel density estimation

```{r echo=TRUE, eval=TRUE}
kde_childcareSG_adaptive <- adaptive.density(childcareSG_ppp.km, method="kernel")
plot(kde_childcareSG_adaptive)
```


#### Compare fixed and adaptive kernel density estimation outputs

```{r echo=TRUE, eval=TRUE}
par(mfrow=c(1,2))
plot(kde_childcareSG.bw, main = "Fixed bandwidth")
plot(kde_childcareSG_adaptive, main = "Adaptive bandwidth")
```

### Converting KDE output into grid object

The result is the same, we just convert it so that it is suitable for mapping purposes

```{r echo=TRUE, eval=TRUE}
gridded_kde_childcareSG_bw <- as.SpatialGridDataFrame.im(kde_childcareSG.bw)
spplot(gridded_kde_childcareSG_bw)
```

#### Converting Gridded Output into Raster 

**raster()** of raster package is used to convert the gridded kernel density objects into RasterLayer object

```{r echo=TRUE, eval=TRUE}
kde_childcareSG_bw_raster <- raster(gridded_kde_childcareSG_bw)
```

#### Properties of kde_childcareSG_bw_raster RasterLayer 

*Note that the CRS property is NA*

```{r echo=TRUE, eval=TRUE}
kde_childcareSG_bw_raster
```

#### Assigning Projection Systems

CRS information will now be included in kde_childcareSG_bw_raster RasterLayer *Note that the crs property is completed*

```{r echo=TRUE, eval=TRUE}
projection(kde_childcareSG_bw_raster) <- CRS("+init=EPSG:3414")
kde_childcareSG_bw_raster
```

### Visualising the output in tmap

```{r echo=TRUE, eval=TRUE}
tm_shape(kde_childcareSG_bw_raster) + 
  tm_raster("v") +
  tm_layout(legend.position = c("right", "bottom"), frame = FALSE)
```

*Note: raster values are encoded explicitly onto the raster pixel using the values in “v” field*

### Comparing Spatial Point Patterns using KDE

KDE of childcare centres will be compared at Ponggol, Tampines, Chua Chu Kang and Jurong West planning areas

#### Extracting Study Area

```{r echo=TRUE, eval=TRUE}
pg = mpsz[mpsz@data$PLN_AREA_N == "PUNGGOL",]
tm = mpsz[mpsz@data$PLN_AREA_N == "TAMPINES",]
ck = mpsz[mpsz@data$PLN_AREA_N == "CHOA CHU KANG",]
jw = mpsz[mpsz@data$PLN_AREA_N == "JURONG WEST",]

par(mfrow=c(2,2))
plot(pg, main = "Ponggol")
plot(tm, main = "Tampines")
plot(ck, main = "Choa Chu Kang")
plot(jw, main = "Jurong West")
```

#### Converting spatial point data frame into generic sp format

Convert SpatialPolygonsDataFrame layers into generic spatialpolygons layers

```{r echo=TRUE, eval=TRUE}
pg_sp = as(pg, "SpatialPolygons")
tm_sp = as(tm, "SpatialPolygons")
ck_sp = as(ck, "SpatialPolygons")
jw_sp = as(jw, "SpatialPolygons")
```

##### 6.1.7.4 Creating Owin Object 

Convert these SpatialPolygons objects into owin objects required by **spatstat**

```{r echo=TRUE, eval=TRUE}
pg_owin = as(pg_sp, "owin")
tm_owin = as(tm_sp, "owin")
ck_owin = as(ck_sp, "owin")
jw_owin = as(jw_sp, "owin")
```

#### Combining childcare points and the study area

Extract childcare centres within the regions for later analysis 

```{r echo=TRUE, eval=TRUE}
childcare_pg_ppp = childcare_ppp_jit[pg_owin]
childcare_tm_ppp = childcare_ppp_jit[tm_owin]
childcare_ck_ppp = childcare_ppp_jit[ck_owin]
childcare_jw_ppp = childcare_ppp_jit[jw_owin]
```

**rescale()** is used to trasnform the unit of measurement from m to km

```{r echo=TRUE, eval=TRUE}
childcare_pg_ppp.km = rescale(childcare_pg_ppp, 1000, "km")
childcare_tm_ppp.km = rescale(childcare_tm_ppp, 1000, "km")
childcare_ck_ppp.km = rescale(childcare_ck_ppp, 1000, "km")
childcare_jw_ppp.km = rescale(childcare_jw_ppp, 1000, "km")

par(mfrow=c(2,2))
plot(childcare_pg_ppp.km, main="Punggol")
plot(childcare_tm_ppp.km, main="Tampines")
plot(childcare_ck_ppp.km, main="Choa Chu Kang")
plot(childcare_jw_ppp.km, main="Jurong West")
```

#### Computing KDE

*bw.diggle* method is used to derive the bandwidth of each KDE

```{r echo=TRUE, eval=TRUE}
par(mfrow=c(2,2))
plot(density(childcare_pg_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tempines")
plot(density(childcare_ck_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Choa Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=bw.diggle, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
```

#### Computing fixed bandwidth KDE

```{r echo=TRUE, eval=TRUE}
par(mfrow=c(2,2))
plot(density(childcare_ck_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Chou Chu Kang")
plot(density(childcare_jw_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Jurong West")
plot(density(childcare_pg_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Punggol")
plot(density(childcare_tm_ppp.km, 
             sigma=0.25, 
             edge=TRUE, 
             kernel="gaussian"),
     main="Tampines")
```

### Nearest Point Analysis

*clarkevans.test()* of statspat is used to perform Clark-Evans test of aggregation for a spatial point pattern 

The test hypotheses are:

- **Ho**: distribution of childcare services are randomly distributed
- **H1**: distribution of childcare services are not randomly distributed

The 95% confident interval will be used

#### Testing spatial point patterns using Clark and Evans Test

```{r echo=TRUE, eval=TRUE}
clarkevans.test(childcareSG_ppp,
                correction="none",
                clipregion="sg_owin",
                alternative=c("clustered"),
                nsim=99)
```

**Conclusion**: Since p-value is less than 0.05, we can therefore reject the null hypothesis that distribution of childcare services are randomly distributed. Since the Nearest Neighbour index is < 1, we can infer that the distribution of childcare services resemble a clustered distribution

#### Clark and Evans Test: Choa Chu Kang planning area

*clarkevans.test()* of spatstat is used to performs Clark-Evans test of aggregation for childcare centres in Choa Chu Kang planning area

```{r echo=TRUE, eval=TRUE}
clarkevans.test(childcare_ck_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

#### Clark and Evans Test: Tampines planning area

```{r echo=TRUE, eval=TRUE}
clarkevans.test(childcare_tm_ppp,
                correction="none",
                clipregion=NULL,
                alternative=c("two.sided"),
                nsim=999)
```

## Spatial Second-order Spatial Point Patterns Analysis

### Analysing Spatial Point Process Using G-Function

G function measures the distribution of the distances from an arbitrary event to its nearest event. *Gest()* and *envelope()* of spatstat() package is used to compute G-function estimation and perform monte carlo simiulation respectively

### Choa Chu Kang planning area

#### Computing G-Function Estimation

*Gest()* of spatstat package is used to compute G function

```{r echo=TRUE, eval=TRUE}
G_CK = Gest(childcare_ck_ppp, correction = "border")
plot(G_CK, xlim=c(0,500))
```
#### Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Choa Chu Kang are randomly distributed
- **H1**: distribution of childcare services at Choa Chu Kang are not randomly distributed

The null hypothesis will be rejected if p-value <  0.001 (alpha value)

Monte Carlo test with G-function

```{r echo=TRUE, eval=TRUE}
G_CK.csr <- envelope(childcare_ck_ppp, Gest, nsim = 999)
```

```{r echo=TRUE, eval=TRUE}
plot(G_CK.csr)
```

### Tampines planning area

#### Computing G-Function Estimation

```{r echo=TRUE, eval=TRUE}
G_tm = Gest(childcare_tm_ppp, correction = "best")
plot(G_tm)
```

#### Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Tampines are randomly distributed
- **H1**: distribution of childcare services at Tampines are not randomly distributed

The null hypothesis will be rejected if p-value <  0.001 (alpha value)

```{r echo=TRUE, eval=TRUE}
G_tm.csr <- envelope(childcare_tm_ppp, Gest, correction = "all", nsim = 999)
```

```{r echo=TRUE, eval=TRUE}
plot(G_tm.csr)
```

### Analysing Spatial Point Process Using F-Function

F function estimates the empty space function F(r) or its hazard rate h(r) from a point pattern in a window of arbitrary shape. *Fest()* and *envelope()* of spatstat() package is used to compute F-function estimation and perform monte carlo simiulation respectively

### Choa Chu Kang planning area

#### Computing F-Function Estimation

*Fest()* of spatstat  package is used to compute F-function

```{r echo=TRUE, eval=TRUE}
F_CK = Fest(childcare_ck_ppp)
plot(F_CK)
```

#### Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Choa Chu Kang are randomly distributed
- **H1**: distribution of childcare services at Choa Chu Kang are not randomly distributed

The null hypothesis will be rejected if p-value < 0.001 (alpha value) 

Monte Carlo test with F-function

```{r echo=TRUE, eval=TRUE}
F_CK.csr <- envelope(childcare_ck_ppp, Fest, nsim = 999)
```

```{r echo=TRUE, eval=TRUE}
plot(F_CK.csr)
```

### Tampines planning area

#### Computing F-Function Estimation

```{r echo=TRUE, eval=TRUE}
F_tm = Fest(childcare_tm_ppp, correction = "best")
plot(F_tm)
```

#### Performing Complete Spatial Randomness Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Tampines are randomly distributed
- **H1**: distribution of childcare services at Tampines are not randomly distributed

The null hypothesis will be rejected if p-value <  0.001 (alpha value)

```{r echo=TRUE, eval=TRUE}
F_tm.csr <- envelope(childcare_tm_ppp, Fest, correction = "all", nsim = 999)
```

```{r echo=TRUE, eval=TRUE}
plot(F_tm.csr)
```

### Analysing Spatial Point Process Using K-Function

K-function measures the number of events found up to a given distance of any particular event. *Kest()* and *envelope()* of spatstat() package is used to compute K-function estimation and perform monte carlo simiulation respectively

### Choa Chu Kang planning area

#### Computing K-Function Estimate

```{r echo=TRUE, eval=TRUE}
K_ck = Kest(childcare_ck_ppp, correction = "Ripley")
plot(K_ck, . -r ~ r, ylab= "K(d)-r", xlab = "d(m)")
```

#### Performing Complete Random Spatial Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Choa Chu Kang are randomly distributed
- **H1**: distribution of childcare services at Choa Chu Kang are not randomly distributed

The null hypothesis will be rejected if p-value < 0.001 (alpha value)

```{r echo=TRUE, eval=TRUE}
K_ck.csr <- envelope(childcare_ck_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE, eval=TRUE}
plot(K_ck.csr, . - r ~ r, xlab="d", ylab="K(d)-r")
```

### Tampines planning area

#### Computing K-Function Estimation

```{r echo=TRUE, eval=TRUE}
K_tm = Kest(childcare_tm_ppp, correction = "Ripley")
plot(K_tm, . -r ~ r, 
     ylab= "K(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### Performing Complete Random Spatial Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Tampines are randomly distributed
- **H1**: distribution of childcare services at Tampines are not randomly distributed

The null hypothesis will be rejected if p-value < 0.001 (alpha value)

```{r echo=TRUE, eval=TRUE}
K_tm.csr <- envelope(childcare_tm_ppp, Kest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE, eval=TRUE}
plot(K_tm.csr, . - r ~ r, 
     xlab="d", ylab="K(d)-r", xlim=c(0,500))
```

### Analysing Spatial Point Process Using L-Function

*Lest()* and *envelope()* of spatstat package is used to compute L-function estimation and perform monte carlo simulation test respectively

### Choa Chu Kang planning area

####Computing L-Function Estimation

```{r echo=TRUE, eval=TRUE}
L_ck = Lest(childcare_ck_ppp, correction = "Ripley")
plot(L_ck, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)")
```

#### Performing Complete Spatial Random Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Choa Chu Kang are randomly distributed
- **H1**: The distribution of childcare services at Choa Chu Kang are not randomly distributed

The null hypothesis will be rejected if p-value < 0.001 (alpha value)

```{r echo=TRUE, eval=TRUE}
L_ck.csr <- envelope(childcare_ck_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE, eval=TRUE}
plot(L_ck.csr, . - r ~ r, xlab="d", ylab="L(d)-r")
```

### Tampines planning area

#### Computing L-Function Estimate

```{r echo=TRUE, eval=TRUE}
L_tm = Lest(childcare_tm_ppp, correction = "Ripley")
plot(L_tm, . -r ~ r, 
     ylab= "L(d)-r", xlab = "d(m)", 
     xlim=c(0,1000))
```

#### Performing Complete Spatial Random Test

To confirm the observed spatial patterns above, a hypothesis test will be conducted:

- **Ho**: distribution of childcare services at Tampines are randomly distributed
- **H1**: distribution of childcare services at Tampines are not randomly distributed

The null hypothesis will be rejected if p-value < 0.001 (alpha value)

```{r echo=TRUE, eval=TRUE}
L_tm.csr <- envelope(childcare_tm_ppp, Lest, nsim = 99, rank = 1, glocal=TRUE)
```

```{r echo=TRUE, eval=TRUE}
plot(L_tm.csr, . - r ~ r, 
     xlab="d", ylab="L(d)-r", xlim=c(0,500))
```