---
title: "In-class Exercise 3"
description: |
  In this in-class exercise, we learn about visualising extreme values through the use of percentile, box and rate maps.
date: 08-30-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1.0 Installing Packages

In this in-class exercise, we will be using the packages sf, tmap and tidyverse respectively. The code chunk below is used to install and load these packages in RStudio.

```{r echo=TRUE, eval=TRUE}
packages = c('sf', 'tmap', 'tidyverse')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## 2.0 The Data
We will be using the dataset which is created from Hands-on Exercise 3. 

### 2.1 Importing and Loading Dataset
The code chunk below uses the read_rds() function of tidyverse package to import the dataset as a simple feature data frame called mpszpop2020.

```{r echo=TRUE, eval=TRUE}
mpszpop2020 = read_rds("data/rds/mpszpop2020.rds")
```

### 2.2 Data Preparation
The mpszpop2020 dataset contains many observations with missing values. As a result, we will have to remove them as shown in the code chunk below.

```{r echo=TRUE, eval=TRUE}
mpszpop2020na <- mpszpop2020 %>% drop_na()
```

## 3.0 Visualising Extreme Values
### 3.1 Boxplot
Boxplot is able to show the statistics and distribution of data values

- Handy in detecting outliers
- Unable to reveal spatial distribution of these outliers


The code chunk below plots a boxplot.

```{r echo=TRUE, eval=TRUE}
ggplot(data=mpszpop2020na, 
       aes(x = "", y = AGED)) + geom_boxplot()
```

### 3.2 Extreme Value Maps

- Variations of choropleth maps where the classification is designed to highlight extreme values at the lower and upper end of the scale, with the goal of identifying outliers

#### 3.2.1 Percentile Map

- Special type of ***quantile*** map with 6 categories: 0-1%, 1-10%, 10-50%, 50-90%, 90-99%, 99-100%
- *Note: The begin and endpoint need to be included.*

##### Extract a variable from a sf dataframe
```{r echo=TRUE, eval=TRUE}
percent <- c(0, .01, .1, .5, .9, .99, 1)
var <- mpszpop2020na["DEPENDENCY"] %>% st_set_geometry(NULL)
quantile(var[,1], percent)
```

##### Advantages of writing functions
- Makes code easier to understand
- As requirements change, we only need to update code in 1 place instead of many
- Can eliminate the chance of making incidental mistakes while copy and pasting

##### Creating the get.var function
```{r echo=TRUE, eval=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

The code chunk above extracts a variable as a vector out of a sf dataframe. 

- Arguments
  - vname: variable name
  - df: name of sf dataframe
- Returns
  - vector with values (without a column name)


##### Plotting a percentile map using tmap functions
```{r echo=TRUE, eval=TRUE}
percent <- c(0, .01, .1, .5, .9, .99, 1)
var <- get.var("DEPENDENCY", mpszpop2020na)
bperc <- quantile(var, percent)
tm_shape(mpszpop2020) + 
  tm_polygons() +
tm_shape(mpszpop2020na) +
  tm_fill("DEPENDENCY", 
          title ="DEPENDENCY", 
          breaks = bperc,
          palette="Blues", 
          labels=c("< 1%", "1% - 10%",
                   "10% - 50%",
                   "50% - 90%",
                   "90% - 99%",
                   "> 99%")) +
  tm_borders() + 
  tm_layout(title = "Percentile Map",
            title.position = c("right", "bottom"))
```

##### Percentile mapping function
```{r echo=TRUE, eval=TRUE}
percentmap <- function(vnam, df, legtitle=NA, mtitle="Percentile Map") {
  percent <- c(0, .01, .1, .5, .9, .99, 1)
  var <- get.var(vnam, df)
  bperc <- quantile(var, percent)
  tm_shape(mpszpop2020) + 
  tm_polygons() + 
  tm_shape(df) +
    tm_fill(vnam,
            title=legtitle,
            breaks=bperc,
            palette="Blues",
            labels=c("<1%", "1-10%", 
                     "10-50%", "50-90%", 
                     "90-99%", ">99%")) + 
  tm_borders() + tm_layout(title=mtitle, title.position = c("right","bottom"))
}
```

##### Testing the percentile mapping function
```{r echo=TRUE, eval=TRUE}
youngmap <- percentmap("YOUNG", mpszpop2020na)
agedmap <- percentmap("AGED", mpszpop2020na)
```

##### "YOUNG" Percent Map 
```{r}
youngmap
```

##### "AGED" Percent Map 
```{r}
agedmap
```

#### 3.2.2 Box Map

- Display summary statistics on a choropleth map by using the basic principles of a boxplot
- Custom breaks specification has to be used to create a box map - however, it varies depending on whether lower or upper outliers are present

##### Creating the boxbreaks function
```{r echo=TRUE, eval=TRUE}
boxbreaks <- function(v,mult=1.5) {
  qv <- unname(quantile(v))
  iqr <- qv[4] - qv[2]
  upfence <- qv[4] + mult * iqr
  lofence <- qv[2] - mult * iqr
  # initialize break points vector
  bb <- vector(mode="numeric",length=7)
  # logic for lower and upper fences
  if (lofence < qv[1]) {  # no lower outliers
    bb[1] <- lofence
    bb[2] <- floor(qv[1])
  } else {
    bb[2] <- lofence
    bb[1] <- qv[1]
  }
  if (upfence > qv[5]) { # no upper outliers
    bb[7] <- upfence
    bb[6] <- ceiling(qv[5])
  } else {
    bb[6] <- upfence
    bb[7] <- qv[5]
  }
  bb[3:5] <- qv[2:4]
  return(bb)
}
```

The code chunk above creates break points for a box map.

- Arguments:
  - v: vecctor with observations
  - mult: multiplier for IQR (*default 1.5*)
- Returns:
  - bb: vector with 7 break points to compute quartiles and fences


##### Creating the get.var function
```{r echo=TRUE, eval=TRUE}
get.var <- function(vname,df) {
  v <- df[vname] %>% st_set_geometry(NULL)
  v <- unname(v[,1])
  return(v)
}
```

The code chunk above is a function to extract a vector out of a sf dataframe.

- Arguments
  - vname: variable name
  - df: name of sf dataframe
- Returns
  - vector with values (without a column name)


##### Test drive the newly created function
```{r echo=TRUE, eval=TRUE}
var <- get.var("AGED", mpszpop2020na) 
boxbreaks(var)
```

##### The code chunk below excludes observations with AGED=NA
```{r echo=TRUE, eval=TRUE}
mpszpop2020a <- mpszpop2020 %>% filter(AGED>=0)
var <- get.var("AGED", mpszpop2020a)
boxbreaks(var)
```

##### Boxmap function
```{r echo=TRUE, eval=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```

The code chunk above is a R function to create a box map. 

- Arguments:
  - vnam: variable name 
  - df: simple features polygon layer
  - legtitle: legend title
  - mtitle: map title
  - mult: multiplier for IQR
- Returns:
  - tmap-element (plots a map)


##### Testing the boxmap function
```{r echo=TRUE, eval=TRUE, fig.width=12, fig.height=6}
boxmap("ECONOMY ACTIVE", mpszpop2020a)
```

- There are 8 upper outliers (i.e. planning subzone with extremely high numbers of economically active population)
- 4 upper outliers are located at the eastern region and they are close to each other
- There is no lower outlier


##### Tidy version of boxmap
```{r echo=TRUE, eval=TRUE}
boxmap <- function(vnam, df, 
                   legtitle=NA,
                   mtitle="Box Map",
                   mult=1.5){
  var <- get.var(vnam,df)
  bb <- boxbreaks(var)
  tm_shape(mpszpop2020) + 
  tm_polygons() +
  tm_shape(df) +
     tm_fill(vnam,title=legtitle,
             breaks=bb,
             palette="Blues",
          labels = c("lower outlier", 
                     "< 25%", 
                     "25% - 50%", 
                     "50% - 75%",
                     "> 75%", 
                     "upper outlier"))  +
  tm_borders() +
  tm_layout(title = mtitle, 
            title.position = c("right",
                               "bottom"))
}
```

##### Testing the tidy version of boxmap function
```{r echo=TRUE, eval=TRUE}
boxmap("ECONOMY ACTIVE", mpszpop2020a)
```

#### 3.2.3 Choropleth Map for Rates
- It is important to account how the population is distributed in space, so that we will not end up mapping population size instead of our topic of interest

##### Computing the raw rate
```{r echo=TRUE, eval=TRUE}
mpszpop2020a <- mpszpop2020 %>% mutate(`AGED%` = (`AGED`/`TOTAL`)*100) %>%
                filter(`AGED%` >= 0)
```

##### Use the boxmap function to plot the raw rate map
```{r echo=TRUE, eval=TRUE}
var <- get.var("AGED%", mpszpop2020a)
boxbreaks(var)
boxmap("AGED%",mpszpop2020a)
```