---
title: "Hands-on Exercise 5"
description: |
  This hands-on exercise covers the appropriate functions of spNetwork package.
author:
  - name: Genice Goh
    url: {}
date: 09-12-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Launching Packages

- spNetwork: spatial point patterns analysis (e.g. KDE, k-function on network)
- rgdal: import geospatial data and store as **Spatial** objects
- sp: provides classes and methods for dealing with spatial data
- tmap: used to plot choropleth maps

```{r echo=TRUE}
packages = c('sp', 'rgdal', 'spNetwork', 'tmap')
for (p in packages){
if(!require(p, character.only = T)){
install.packages(p)
}
library(p,character.only = T)
}
```

## Data Import and Preparation

*readOGD()* of rgdal package is used to important Punggol_St and Punggol_CC geospatial data sets as SpatialLinesDataFrame and SpatialPointsDataFrame respectively

```{r echo=TRUE}
network <- readOGR(dsn="data", 
                   layer="Punggol_St",
                   verbose = FALSE)

childcare <- readOGR(dsn="data",
                     layer="Punggol_CC",
                     verbose = FALSE)
```

<br>

*str()* is used to examine the structure of the output Spatial DataFrame 

```{r echo=TRUE}
str(network)
str(childcare)
```

<br>

*spNetwork* expects the geospatial data to contain complete CRS information. *spTranform()* of sp package is used to assign EPSG code to the SpatialDataFrames

```{r echo=TRUE}
childcare <-spTransform(childcare, CRS("+init=epsg:3414"))
network <- spTransform(network,CRS("+init=epsg:3414"))
```

## Visualising Geospatial Data

Good practice to visualise geospatial data: using plot() / tmap package (to visualise geospatial data with high cartographic quality and interactive manner) 

```{r echo=TRUE}
plot(network)
plot(childcare,add=T,col='red',pch = 19)
```

```{r echo=TRUE}
tmap_mode('view')
tm_shape(childcare)+
  tm_dots() +
tm_shape(network)+
  tm_lines()
```

```{r echo=TRUE}
tmap_mode('plot')
```

## Network Constrained KDE (NetKDE) Analysis

Appropriate functions from the *spNetwork* package will be used to perform NetKDE analysis

### Preparing lixels objects

*lixelize_lines()* of spNetwork is used to cut the SpatialLines object into lixels with a specified minimal distance. This needs to be done before computing NetKDE

```{r echo=TRUE}
lixels <- lixelize_lines(network,700,mindist = 350)
```

- length of a lixel *lx_length* is set to 700m
- minimum length of a lixel *mindist* is set to 350m

If the length of the final lixel < minimum distance after cutting, then it is added to the previous lixel. If NULL, then mindist = maxdist/10. Also, segments that are already shorter than the minimum distance are not modified

*Note: There is another function called lixelize_lines.mc() which provides multicore support*

### Generating line centre points

*lines_center()* of spNetwork is used to generate a SpatialPointsDataFrame with line centre points (i.e. samples)

- Points are located at the centre of the line based on the line length

```{r echo=TRUE}
samples <- lines_center(lixels)
```

### Performing NetKDE

```{r echo=TRUE}
densities <- nkde(network, 
                  events = childcare,
                  w = rep(1,nrow(childcare)),
                  samples = samples,
                  kernel_name = "quartic",
                  bw = 300, 
                  div= "bw", 
                  method = "simple", 
                  digits = 1, 
                  tol = 1,
                  grid_shape = c(1,1), 
                  max_depth = 8,
                  agg = 5, #we aggregate events within a 5m radius (faster calculation)
                  sparse = TRUE,
                  verbose = FALSE)
```

- kernel_name: quartic kernel is used. Possible kernel methods supported by spNetwork: *triangle, gaussian, scaled gaussian, tricube, cosine ,triweight, epanechnikov or uniform*
- method: simple method is used to calculate NKD. *spNetwork* currently supports three methods are:
  - simple: the distances between events and sampling points are replaced by network distances, and the formula of the kernel is adapted to calculate the density over a linear unit instead of an areal unit
  - discontinuous: equally “divides” the mass density of an event at intersections of lixels
  - continuous: divide the mass of the density at intersection but adjusts the density before the intersection to make the function continuous
  
### Visualising NetKDE

Computed density values need to be inserted into samples and lixels objects as *densities* field before visualising NetKDE values

```{r echo=TRUE}
samples$density <- densities
lixels$density <- densities
```

Since SVY21 projection system is in m, the computed density values are very small. The density values are therefore rescaled from number of events/m to number of events/km

```{r echo=TRUE}
samples$density <- samples$density*1000
lixels$density <- lixels$density*1000
```

```{r echo=TRUE}
tmap_mode('view')
tm_shape(lixels)+
  tm_lines(col="density")+
tm_shape(childcare)+
  tm_dots()
```

```{r echo=TRUE}
tmap_mode('plot')
```

The interactive map reveals road segments with relatively higher density of childcare centres (darker color) than road segments with relatively lower density of childcare centres (lighter color)

## Network Constrained G and K-Function Analysis

*kfunctions()* of spNetwork package is used to perform complete spatial randomness (CSR) test by using kfunctions() of spNetwork package

**Ho**: The observed spatial point events (i.e distribution of childcare centres) are uniformly distributed over a street network in Punggol Planning Area

The CSR test is based on the assumption of the binomial point process which implies the hypothesis that the childcare centres are **randomly and independently distributed** over the street network

If hypothesis is rejected, we may infer that the distribution of childcare centres are **spatially interacting and dependent on each other**; as a result, they may **form nonrandom patterns**

```{r echo=TRUE}
kfun_childcare <- kfunctions(network, 
                             childcare,
                             start = 0, 
                             end = 1000, 
                             step = 50, 
                             width = 50, 
                             nsim = 50, 
                             resolution = 50,
                             verbose = FALSE, 
                             conf_int = 0.05)
```

- lines: SpatialLinesDataFrame with the sampling points. Geometries must be a SpatialLinesDataFrame (may crash if some geometries are invalid)
- points: SpatialPointsDataFrame representing the points on the network. These points will be snapped on the network
- start: start value (double) for evaluating the k and g functions
- end: last value (double) for evaluating the k and g functions
- step: A jump (double) between two evaluations of the k and g function
- width: width of each donut for the g-function
- nsim: integer indicating number of Monte Carlo simulations required. *Note: most of the time, more simulations are required for inference*
- resolution: Selecting a resolution will greatly reduce calculation time when simulating random points on the network. When resolution is null, the random points can occur everywhere on the graph. If a value is specified, the edges are split according to it and the random points are selected vertices on the new network
- conf_int: double indicating the width confidence interval (default = 0.05)


The output of kfunctions() is a list with the following values:

- plotkA: ggplot2 object representing values of k-function
- plotgA: ggplot2 object representing values of g-function
- valuesA: DataFrame with values used to build the plots

### Visualise K-function ggplot2 object

```{r echo=TRUE}
kfun_childcare$plotk  
```

The blue line is the empirical network K-function of the childcare centres in Punggol planning area. The gray envelop represents the results of the 50 simulations in the interval 2.5% - 97.5%. Because the blue line between the distance of 250m-400m are below the gray area, we can infer that the childcare centres in Punggol planning area resemble regular pattern at the distance of 250m-400m