---
title: "Hands-on Exercise 8"
description: |
  In this hands-on exercise, we learn how to perform geographical segmentation by using appropriate R packages. We will also be exposed to R packages for performing cluster analysis and visualising clustering results.
author:
  - name: Genice Goh
    url: {}
date: 10-10-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, cache = TRUE)
```

## The Analytical Question

In geobusiness and spatial policy, it is common practice to delineate the market or planning area into homogeneous regions by using multivariate data. 

In this hands-on exercise, we are interested to delineate Shan State, Myanmar into homogeneous regions by using multiple Information and Communication technology (ICT) measures, namely: radio, television, land line phone, mobile phone, computer, and Internet at home.

## Installing and Loading Packages

The following packages will be used for this analysis:

- **sf**, **rgdal**, **spdep**: used for handling of geospatial data
- **tidyverse** (**readr**, **ggplot2**, **dplyr**): mainly used for wrangling attribute data 
- **tmap**: used to prepare cartographic quality chropleth maps
- **coorplot**, **ggpubr**, **heatmaply**: used for multivariate data visualisation and analysis
- **cluster**, **ClustGeo**: used to perform cluster analysis


```{r echo=TRUE}
packages = c('rgdal', 'spdep', 'tmap', 'sf', 'ggpubr', 'cluster', 'ClustGeo', 'factoextra', 'NbClust', 'heatmaply', 'corrplot', 'psych', 'tidyverse')

for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
    }
  library(p,character.only = T)
}
```

## Importing Geospatial Data

*st_read()* of **sf** package is used to import the `Myanmar Township Boundary GIS data` shapefile.

```{r echo=TRUE}
shan_sf <- st_read(dsn = "data/geospatial", 
                   layer = "myanmar_township_boundaries") %>%
  filter(ST %in% c("Shan (East)", "Shan (North)", "Shan (South)"))
```

The imported township boundary object is called `shan_sf`. It is saved in **simple feature data.frame** format.

```{r echo=TRUE}
shan_sf
```

Notice that sf.data.frame is conformed to Hardy Wickham’s tidy framework. Since `shan_sf` is conformed to tidy framework, we can also *glimpse()* to reveal the data type of it’s fields.

```{r echo=TRUE}
glimpse(shan_sf)
```

## Importing Aspatial Data and Data Wrangling

We will import Shan-ICT.csv using *read_csv()* of **readr package**. 

The imported InfoComm variables are extracted from **The 2014 Myanmar Population and Housing Census Myanmar**. It is saved in R’s **tibble data.frame** format.

```{r echo=TRUE}
ict <- read_csv ("data/aspatial/Shan-ICT.csv")
summary(ict)
```

There are a total of **11 fields** and **55 observations** in the **tibble data.frame**.

### Derive new variables

The unit of measurement of the variables is the **number of households**. Using these values directly will be biased due to the underlying total number of households. In general, townships with higher total number of households will also have higher number of households owning radio, TV, etc.

In order to overcome this problem, we will derive the penetration rate of each ICT variable.

```{r echo=TRUE}
ict_derived <- ict %>%
  mutate(`RADIO_PR` = `Radio`/`Total households`*1000) %>%
  mutate(`TV_PR` = `Television`/`Total households`*1000) %>%
  mutate(`LLPHONE_PR` = `Land line phone`/`Total households`*1000) %>%
  mutate(`MPHONE_PR` = `Mobile phone`/`Total households`*1000) %>%
  mutate(`COMPUTER_PR` = `Computer`/`Total households`*1000) %>%
  mutate(`INTERNET_PR` = `Internet at home`/`Total households`*1000) %>%
  rename(`DT_PCODE` =`District Pcode`,`DT`=`District Name`,
         `TS_PCODE`=`Township Pcode`, `TS`=`Township Name`,
         `TT_HOUSEHOLDS`=`Total households`,
         `RADIO`=`Radio`, `TV`=`Television`, 
         `LLPHONE`=`Land line phone`, `MPHONE`=`Mobile phone`,
         `COMPUTER`=`Computer`, `INTERNET`=`Internet at home`) 
```

```{r echo=TRUE}
summary(ict_derived)
```

Notice that six new fields have been added into the data.frame: `RADIO_PR`, `TV_PR`, `LLPHONE_PR`, `MPHONE_PR`, `COMPUTER_PR` and `INTERNET_PR`.

## Exploratory Data Analysis (EDA)

### EDA using statistical graphs

We can plot the distribution of the variables by using the appropriate EDA techniques.

**Histogram** is useful to identify the **overall distribution** of the data values (i.e. left skew, right skew or normal distribution).

```{r echo=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

**Boxplot** is useful to detect if there are **outliers**.

```{r echo=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO`)) +
  geom_boxplot(color="black", fill="light blue")
```

Next, we will also be plotting the distribution of the newly derived variables (i.e. radio penetration rate).

```{r echo=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, color="black", fill="light blue")
```

```{r echo=TRUE}
ggplot(data=ict_derived, aes(x=`RADIO_PR`)) +
  geom_boxplot(color="black", fill="light blue")
```

What can be observed from the histogram and boxplot?

- The distribution of radio penetration rates is slightly **right skewed**.
- The **median** radio penetration rate is approximately 210, and there is **1 outlier** with a value of approximately 480.

We can also plot multiple histograms to reveal the distribution of the selected variables in the `ict_derived` data.frame. We will first create the individual histograms before grouping them together using *ggarange()* of **ggpubr** package

```{r echo=TRUE}
radio <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

tv <- ggplot(data=ict_derived, 
             aes(x= `TV_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

llphone <- ggplot(data=ict_derived, 
             aes(x= `LLPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

mphone <- ggplot(data=ict_derived, 
             aes(x= `MPHONE_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

computer <- ggplot(data=ict_derived, 
             aes(x= `COMPUTER_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

internet <- ggplot(data=ict_derived, 
             aes(x= `INTERNET_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

ggarrange(radio, tv, llphone, mphone, computer, internet, 
          ncol = 3, 
          nrow = 2)
```

### EDA using choropleth maps

### Joining geospatial data with aspatial data

Before we can prepare the choropleth map, we need to combine both the geospatial data object and aspatial data.frame object. This will be performed using *left_join()* of **dplyr** package. 

The `shan_sf` simple feature data.frame will be used as the base data object and the `ict_derived` data.frame will be used as the join table. The unique identifier used to join both data objects is `TS_PCODE`.

```{r echo=TRUE}
shan_sf <- left_join(shan_sf, ict_derived, by=c("TS_PCODE"="TS_PCODE"))
```

It is important to note that no new output data has been created. Instead, the data fields from `ict_derived` data frame are now updated into the data frame of `shan_sf`.

### Preparing a choropleth map

To have a quick look at the distribution of radio penetration rate of Shan State at township level, a choropleth map will be prepared using *qtm()* of **tmap** package.

```{r echo=TRUE}
qtm(shan_sf, "RADIO_PR")
```

In order to reveal that the number of households with radio is biased to the underlying total number of households at the township level, we will plot these two choropleth maps side by side.

```{r echo=TRUE, fig.width=10, fig.height=8}
TT_HOUSEHOLDS.map <- tm_shape(shan_sf) + 
  tm_fill(col = "TT_HOUSEHOLDS",
          n = 5,
          style = "jenks", 
          title = "Total households") + 
  tm_borders(alpha = 0.5) 

RADIO.map <- tm_shape(shan_sf) + 
  tm_fill(col = "RADIO",
          n = 5,
          style = "jenks",
          title = "Households with Radio ") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(TT_HOUSEHOLDS.map, RADIO.map,
             asp=NA, ncol=2)
```

Notice that the choropleth maps above clearly show that townships with relatively larger number of households are also showing relatively higher number of radio ownership.

Let us plot the choropleth maps showing the distributions of total number of households and radio penetration rate.

```{r echo=TRUE, fig.width=8, fig.height=5}
tm_shape(shan_sf) +
    tm_polygons(c("TT_HOUSEHOLDS", "RADIO_PR"),
                style="jenks") +
    tm_facets(sync = TRUE, ncol = 2) +
  tm_legend(legend.position = c("right", "bottom"))+
  tm_layout(outer.margins=0, asp=0)
```

We can now see that townships with relatively larger number of households do not necessarily have relatively higher radio penetration rates.

## Correlation Analysis

Before we perform cluster analysis, it is important for us to ensure that the cluster variables are not highly correlated.

*corrplot.mixed()* of **corrplot** package is used to visualise and analyse the correlation of the input variables.

```{r echo=TRUE}
cluster_vars.cor = cor(ict_derived[,12:17])
corrplot.mixed(cluster_vars.cor,
         lower = "ellipse", 
               upper = "number",
               tl.pos = "lt",
               diag = "l",
               tl.col = "black")
```

The correlation plot above shows that `COMPUTER_PR` and `INTERNET_PR` are highly correlated. This suggests that only one of them should be used in the cluster analysis instead of both.

## Hierarchical Cluster Analysis

### Extracting clustering variables

The code chunk below will be used to extract the clustering variables from the `shan_sf` simple feature object into data.frame.

```{r echo=TRUE}
cluster_vars <- shan_sf %>%
  st_set_geometry(NULL) %>%
  select("TS.x", "RADIO_PR", "TV_PR", "LLPHONE_PR", "MPHONE_PR", "COMPUTER_PR")
head(cluster_vars,10)
```

Notice that the final clustering variables list does not include variable `INTERNET_PR` because it is highly correlated with variable `COMPUTER_PR`.

Next, we need to change the rows by township name instead of row number.

```{r echo=TRUE}
row.names(cluster_vars) <- cluster_vars$"TS.x"
head(cluster_vars,10)
```

We can now delete the `TS.x` field.

```{r echo=TRUE}
shan_ict <- select(cluster_vars, c(2:6))
head(shan_ict, 10)
```

### Data Standardisation

In general, multiple variables will be used in cluster analysis. It is not unusual their values range are different. In order to prevent the cluster analysis result from being biased to clustering variables with large values, it is useful to **standardise** the input variables before performing cluster analysis.

#### Min-Max Standardisation

*normalize()* of **heatmaply** package is used to standardise the clustering variables using Min-Max method. *summary()* is then used to display the summary statistics of the standardised variables.

```{r echo=TRUE}
shan_ict.std <- normalize(shan_ict)
summary(shan_ict.std)
```

Notice that the values of the Min-max standardised variables now **range from 0 to 1**.

#### Z-score Standardisation

Z-score standardisation of the clustering variables can be easily performed using *scale()* of **Base R**. 

Note: *describe()* of **psych** package is used instead of *summary()* of **Base R** because the earlier provides the **standard deviation**.

```{r echo=TRUE}
shan_ict.z <- scale(shan_ict)
describe(shan_ict.z)
```

Notice that the **mean** and **standard deviation** of the Z-score standardised variables are **0** and **1** respectively.

**Warning**: Z-score standardisation method should only be used if we would assume **all variables come from some normal distribution**.

#### Visualising the standardised clustering variables

Besides reviewing the summary statistics of the standardised clustering variables, it is also a good practice to visualise their distributions graphically.

```{r echo=TRUE, fig.width=8}
r <- ggplot(data=ict_derived, 
             aes(x= `RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue")

shan_ict_s_df <- as.data.frame(shan_ict.std)
s <- ggplot(data=shan_ict_s_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Min-Max Standardisation")

shan_ict_z_df <- as.data.frame(shan_ict.z)
z <- ggplot(data=shan_ict_z_df, 
       aes(x=`RADIO_PR`)) +
  geom_histogram(bins=20, 
                 color="black", 
                 fill="light blue") +
  ggtitle("Z-score Standardisation")

ggarrange(r, s, z,
          ncol = 3,
          nrow = 1)
```

Notice that the overall distribution of the clustering variables will change after data standardisation. Hence, it is advisable **NOT** to perform data standardisation if the **values range** of the clustering variables are **not very large**.

## Computing proximity matrix

Many packages provide functions to calculate distance matrix in R. We will compute the proximity matrix by using *dist()* of **R**.

*dist()* supports six distance proximity calculations: **euclidean (default)**, **maximum**, **manhattan**, **canberra**, **binary** and **minkowski**.

```{r echo=TRUE}
proxmat <- dist(shan_ict, method = 'euclidean')
proxmat
```
## Computing hierarchical clustering

There are several packages that provide hierarchical clustering functions in R. In this hands-on exercise, *hclust()* of **R stats** will be used.

*hclust()* employs **agglomeration** method to compute the cluster. 8 clustering algorithms are supported: **ward.D**, **ward.D2**, **single**, **complete**, **average(UPGMA)**, **mcquitty(WPGMA)**, **median(WPGMC)** and **centroid(UPGMC)**.

The code chunk below performs hierarchical clustering analysis using **ward.D** method. The hierarchical clustering output is stored in an object of class **hclust** which describes the tree produced by the clustering process.

```{r echo=TRUE}
hclust_ward <- hclust(proxmat, method = 'ward.D')
```

*plot()* of **R Graphics** is then used to plot the tree.

```{r echo=TRUE}
plot(hclust_ward, cex = 0.6)
```

## Selecting the optimal clustering algorithm

One of the challenges in performing hierarchical clustering is to identify **stronger clustering structures**. The issue can be solved by using *agnes()* of **cluster** package. 

It functions like *hclust()*, however, with *agnes()* you can also get the **agglomerative coefficient**, which measures the **amount of clustering structure** found (values closer to 1 suggest strong clustering structure).

The agglomerative coefficients of all hierarchical clustering algorithms will now be computed.

```{r echo=TRUE}
m <- c( "average", "single", "complete", "ward")
names(m) <- c( "average", "single", "complete", "ward")

ac <- function(x) {
  agnes(shan_ict, method = x)$ac
}

map_dbl(m, ac)
```

With reference to the output above, we can see that **Ward**’s method provides the **strongest clustering structure** among the four methods assessed. Hence, in the subsequent analysis, only Ward’s method will be used.

## Determining Optimal Clusters

Another technical challenge faced in performing clustering analysis is to determine the **optimal clusters to retain**.

There are 3 commonly used methods to determine the optimal clusters:

- Elbow Method
- Average Silhouette Method
- Gap Statistic Method

### Gap Statistic Method

The gap statistic compares the total within intra-cluster variation for different values of k with their expected values under null reference distribution of the data. The estimate of the optimal clusters will be the **value** that yields **the largest gap statistic**. This means that the clustering structure is far away from the random uniform distribution of points.

To compute the gap statistic, *clusGap()* of **cluster** package will be used. Note that *hcut()* is from the **factoextra** package.

```{r echo=TRUE}
set.seed(12345)
gap_stat <- clusGap(shan_ict, FUN = hcut, nstart = 25, K.max = 10, B = 50)
# Print the result
print(gap_stat, method = "firstmax")
```

Next, we can visualise the plot using *fviz_gap_stat()* of **factoextra** package.

```{r echo=TRUE}
fviz_gap_stat(gap_stat)
```

With reference to the gap statistic graph above, the recommended number of cluster to retain is 1. However, it is **not logical** to retain only one cluster. By examining the gap statistic graph, the **6-cluster** gives the **largest gap statistic** and should be the next best cluster to pick.

Note: In addition to these commonly used approaches, the **NbClust** package provides 30 indices for determining the relevant number of clusters and proposes to users the best clustering scheme from the different results obtained by varying all combinations of number of clusters, distance measures, and clustering methods.

## Interpreting the dendrograms

In the dendrogram displayed above, **each leaf** corresponds to **one observation**. As we move up the tree, **similar observations** are **combined into branches**, which are themselves fused at a higher height.

The height of the fusion, provided on the vertical axis, indicates the **(dis)similarity between two observations**. The **higher** the height of the fusion, the **less similar** the observations are. 

Note: Conclusions about the **proximity of two observations** can be drawn only based on the **height** where **branches containing those two observations first are fused**. We cannot use the proximity of two observations along the horizontal axis as a criteria of their similarity.

It’s also possible to draw the dendrogram with a **border around the selected clusters** by using *rect.hclust()* of **R stats**. The argument border is used to specify the border colors for the rectangles.

```{r echo=TRUE}
plot(hclust_ward, cex = 0.6)
rect.hclust(hclust_ward, k = 6, border = 2:5)
```

## Visually-driven hierarchical clustering analysis

In this section, we will learn how to perform visually-driven hiearchical clustering analysis by using the **heatmaply** package.

With **heatmaply**, we are able to build both highly interactive cluster heatmaps and static cluster heatmaps.

### Transforming the data frame into a matrix

The data was loaded into a data frame, but it has to be a data matrix to create the heatmap.

```{r echo=TRUE}
shan_ict_mat <- data.matrix(shan_ict)
```

### Plotting interactive cluster heatmap using heatmaply()

*heatmaply()* is used to build an interactive cluster heatmap.

```{r echo=TRUE}
heatmaply(normalize(shan_ict_mat),
          Colv=NA,
          dist_method = "euclidean",
          hclust_method = "ward.D",
          seriate = "OLO",
          colors = Blues,
          k_row = 6,
          margins = c(NA,200,60,NA),
          fontsize_row = 4,
          fontsize_col = 5,
          main="Geographic Segmentation of Shan State by ICT indicators",
          xlab = "ICT Indicators",
          ylab = "Townships of Shan State"
          )

```

## Mapping the clusters formed

After close examination of the dendrogram above, we have decided to retain 6 clusters.

*cutree()* of **R Base** will be used to derive a 6-cluster model. The output `groups` is a list object.

```{r echo=TRUE}
groups <- as.factor(cutree(hclust_ward, k=6))
```

In order to visualise the clusters, the `groups` object need to be appended onto `shan_sf` simple feature object.

- The `groups` list object will first be converted into a matrix.
- *cbind()* is then used to append `groups` matrix onto `shan_sf` to produce an output simple feature object called `shan_sf_cluster`.
- *rename()* of **dplyr** package is lastly used to rename `as.matrix.groups` field as `CLUSTER`.

```{r echo=TRUE}
shan_sf_cluster <- cbind(shan_sf, as.matrix(groups)) %>%
  rename(`CLUSTER`=`as.matrix.groups.`)
```

Next, *qtm()* of **tmap** package is used to plot the choropleth map showing the clusters formed.

```{r echo=TRUE}
qtm(shan_sf_cluster, "CLUSTER")
```

The choropleth map above reveals the clusters are **very fragmented**. This ia one of the **major limitation** when non-spatial clustering algorithm such as hierarchical clustering analysis method is used.

## Spatially Constrained Clustering - SKATER approach

In this section, we learn how to derive spatially constrained clusters using the SKATER method.

### Converting into SpatialPolygonsDataFrame

Firstly, we need to convert `shan_sf` into **SpatialPolygonDataFrame**. This is because the SKATER function only supports **sp** objects such as **SpatialPolygonDataFrame**.

*as_Spatial()* of **sf** package to convert `shan_sf` into a **SpatialPolygonDataFrame** called `shan_sp`.

```{r echo=TRUE}
shan_sp <- as_Spatial(shan_sf)
```

## Computing Neighbour List

Next, *poly2nd()* of **spdep** package is used to compute the neighbours list from polygon list.

```{r echo=TRUE}
shan.nb <- poly2nb(shan_sp)
summary(shan.nb)
```

We can plot the neighbours list on `shan_sp`. The **first** plot command gives the **boundaries**. This is followed by the **plot of the neighbor list object**, with coordinates applied to the original SpatialPolygonDataFrame to extract the centroids of the polygons. These are used as the nodes for the graph representation. We also set the color to blue and specify add=TRUE to plot the network on top of the boundaries.

```{r echo=TRUE, fig.width=8, fig.height=6}
plot(shan_sp, border=grey(.5))
plot(shan.nb, coordinates(shan_sp), col="blue", add=TRUE)
```

Note: If you plot the network first and then the boundaries, some of the areas will be clipped. This is because the plotting area is determined by the characteristics of the first plot. In this example, because the boundary map extends further than the graph, we plot it first.

## Computing minimum spanning tree

### Calculating edge costs

Next, *nbcosts()* of **spdep** package is used to compute the cost of each edge (distance between the nodes). This function compute this distance using a data.frame with observations vector in each node.

```{r echo=TRUE}
lcosts <- nbcosts(shan.nb, shan_ict)
```

For each observation, this gives the pairwise dissimilarity between its values on the five variables and the values for the neighbouring observation (from the neighbour list). Basically, this is the notion of a generalised weight for a spatial weights matrix.

Next, we will incorporate these costs into a weights object in the same way as we did in the calculation of inverse of distance weights. In other words, we convert the neighbour list to a list weights object by specifying the just computed **lcosts** as the weights.

In order to achieve this, *nb2listw()* of **spdep** package is used.

Note that we specify the style as **B** to make sure the cost values are not row-standardised.

```{r echo=TRUE}
shan.w <- nb2listw(shan.nb, lcosts, style="B")
summary(shan.w)
```

### Computing minimum spanning tree

The minimum spanning tree is computed using *mstree()* of **spdep** package.

```{r echo=TRUE}
shan.mst <- mstree(shan.w)
```

After computing the MST, we can check its **class** and **dimension**.

```{r echo=TRUE}
class(shan.mst)
dim(shan.mst)
```
Note that the dimension is **54** and not 55. This is because the minimum spanning tree consists of **n-1 edges** (links) in order to traverse all the nodes.

We can display the content of `shan.mst` using *head()*.

```{r echo=TRUE}
head(shan.mst)
```

The plot method for the MST includes a way to show the observation numbers of the nodes in addition to the edge. Similarly, we will plot this together with the township boundaries. We can see how the initial neighbour list is simplified to just one edge connecting each of the nodes, while passing through all the nodes.

```{r echo=TRUE, fig.width=8, fig.height=6}
plot(shan_sp, border=gray(.5))
plot.mst(shan.mst, coordinates(shan_sp), 
     col="blue", cex.lab=0.7, cex.circles=0.005, add=TRUE)
```

## Computing spatially constrained clusters using SKATER method

*skater()* of **spdep** package is used to compute the spatially constrained cluster.

```{r echo=TRUE}
clust6 <- skater(shan.mst[,1:2], shan_ict, method = "euclidean", 5)
```

The skater() takes three mandatory arguments: 

- The first two columns of the MST matrix (i.e. not the cost)
- The data matrix (to update the costs as units are being grouped)
- The number of cuts 
  - Note: It is set to **one less than the number of clusters**. The value specified is **not** the number of clusters, but the number of cuts in the graph.

The result of the *skater()* is an object of class **skater**.

```{r echo=TRUE}
str(clust6)
```

The most interesting component of this list structure is the groups vector containing the labels of the cluster to which each observation belongs (as before, the label itself is arbitary). 

This is followed by a detailed summary for each of the clusters in the edges.groups list. Sum of squares measures are given as ssto for the total and ssw to show the effect of each of the cuts on the overall criterion.

We can check the cluster assignment.

```{r echo=TRUE}
ccs6 <- clust6$groups
ccs6
```

We can find out how many observations are in each cluster using *table()*. We can also find this as the dimension of each vector in the lists contained in edges.groups. For example, the first list has node with dimension 12, which is also the number of observations in the first cluster.

```{r echo=TRUE}
table(ccs6)
```

Lastly, we can also plot the pruned tree that shows the five clusters on top of the townshop area.

```{r echo=TRUE, fig.width=8, fig.height=6}
plot(shan_sp, border=gray(.5))
plot(clust6, coordinates(shan_sp), cex.lab=.7,
     groups.colors=c("red","green","blue", "brown", "pink"), cex.circles=0.005, add=TRUE)
```

## Visualising the clusters in choropleth map

The code chunk below is used to plot the newly derived clusters by using SKATER method.

```{r echo=TRUE}
groups_mat <- as.matrix(clust6$groups)
shan_sf_spatialcluster <- cbind(shan_sf_cluster, as.factor(groups_mat)) %>%
  rename(`SP_CLUSTER`=`as.factor.groups_mat.`)
qtm(shan_sf_spatialcluster, "SP_CLUSTER")
```
For easy comparison, it will be better to place both the hierarchical clustering and spatially constrained hierarchical clustering maps next to each other.

```{r echo=TRUE, fig.width=8, fig.height=6}
hclust.map <- qtm(shan_sf_cluster,
                  "CLUSTER") + 
  tm_borders(alpha = 0.5) 

shclust.map <- qtm(shan_sf_spatialcluster,
                   "SP_CLUSTER") + 
  tm_borders(alpha = 0.5) 

tmap_arrange(hclust.map, shclust.map,
             asp=NA, ncol=2)
```