---
title: "Hands-on Exercise 10"
description: |
  In this hands-on exercise, we learn how to calibrate spatial interaction models by using GLM() of Base R.
author:
  - name: Genice Goh
    url: {}
date: 10-23-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Introduction

We will gain hands-on experience on how to calibrate Spatial Interaction Models (SIM) by using GLM() of Base R. The use case is adapted from Modelling population flows using spatial interaction models by Adam Dennett.

## Installing and Loading Packages

The following packages will be used for this analysis:

- **sf**, **sp**, **geojsonio**, **stplanr**: used for handling of geospatial data
- **tidyverse** (**readr**, **dplyr**, **reshape2**): mainly used for wrangling attribute data 
- **tmap**: used to prepare cartographic quality choropleth maps
- **ggplot2**: used for plotting statistical graphs
- **caret**: used for statistical analysis

```{r echo=TRUE}
packages = c('tmap', 'tidyverse', 'sp', 'caret', 'geojsonio', 'stplanr',
             'reshape2')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

Due to the issue regarding the s2 object class, we will use the older version (i.e. 0.9-8) of **sf** package instead of the latest one (i.e. 1.0-3). We will only need to install once.

```{r echo=TRUE}
library(devtools)

install_version("sf", version= "0.9-8", repos= "http://cran.us.r-project.org")
```

After installation, we will need to launch the library.

```{r echo=TRUE}
library(sf)
```

# Geospatial Data

We will download a copy of the Greater Capital City Statistical Areas boundary layer from a dropbox depository using *geojson_read()* of **geojsonio** package.

```{r echo=TRUE}
Aus <- geojson_read("https://www.dropbox.com/s/0fg80nzcxcsybii/GCCSA_2016_AUST_New.geojson?raw=1", what = "sp")
```

We will extract the data next.

```{r echo=TRUE}
Ausdata <- Aus@data
```

The original data is in **geojson** format. We will convert it into a **simple features** object and set its coordinate reference system in case the file doesn’t have one.

```{r echo=TRUE}
AusSF <- st_as_sf(Aus) %>% 
  st_set_crs(4283) 
```

We will next check if all the simple features are valid.

```{r echo=TRUE}
st_is_valid(AusSF)
```

We can observe from the output message that there are several invalid features. We will therefore have to fix them.

```{r echo=TRUE}
AusSF <- st_make_valid(AusSF)
```

```{r echo=TRUE}
st_is_valid(AusSF)
```
### Displaying the boundary layer

It is good practice to plot the data and check if the boundary layer is correct. *qtm()* of **tmap** package is used to plot `AusSF` simple feature data frame.

```{r echo=TRUE}
tmap_mode("plot")
qtm(AusSF)
```

### Displaying data table

```{r echo=TRUE}
head(AusSF, 10)
```

With close examination, we can notice that the code order is a bit weird. We can fix it by reordering.

```{r echo=TRUE}
AusSF1 <- AusSF[order(AusSF$GCCSA_CODE),]
```

We can now take a look at the data frame again.

```{r echo=TRUE}
head(AusSF1, 10)
```

### Converting into sp object

We will convert the newly ordered SF1 data.frame into an ‘sp’ object. from our.

```{r echo=TRUE}
Aus <- as(AusSF1, "Spatial")
```

## Calculating a distance matrix

In spatial interaction models, space is one of the key predictor variables. In this example, we will use the Euclidean distance measure between the centroids of the Greater Capital City Statistical Areas as our measure of space.

**Caution note: With some areas so huge, there are obvious potential issues (e.g. we could use the average distance to larger settlements in the non-city areas). However, we will proceed with a simple solution for now since this is just an example.**

### Re-projecting to projected coordinate system

The original data is in geographical coordinate system and the unit of measurement is in decimal degree, which is not appropriate for distance measurement. Before we compute the distance matrix, we will re-project the Aus into projected coordinate system using *spTransform()* of **sp** package.

```{r echo=TRUE}
AusProj <- spTransform(Aus,"+init=epsg:3112")
summary(AusProj)
```

### Computing distance matrix

We can technically use *st_distance()* of **sf** package to compute the distance matrix. However, in view of the longer processing time, *spDist()* of **sp** package is used.

```{r echo=TRUE}
dist <- spDists(AusProj)
dist 
```

### Converting distance matrix into distance pair list

In order to integrate the distance matrix with the migration flow data frame later, we need to transform the newly derived distance matrix into a 3-column distance values list.

*melt()* of **reshape2** package is used to complete the task. We can also achieve the same task using *pivot_longer()* of **dplyr** package.

```{r echo=TRUE}
distPair <- melt(dist)
head(distPair, 10)
```

### Converting unit of measurement from m into km

The unit of measurement of Australia projected coordinate system is in **metres**. As a result, the values in the distance matrix are in **metres** too. We will now convert the distance values into **kilometres**.

```{r echo=TRUE}
distPair$value <- distPair$value / 1000
head(distPair, 10)
```

# Importing Interaction Data

We will import the migration data using *read_csv()*of **readr** package.

```{r echo=TRUE}
mdata <- read_csv("https://www.dropbox.com/s/wi3zxlq5pff1yda/AusMig2011.csv?raw=1",col_names = TRUE)

glimpse(mdata)
```

## Combining the imported migration data

We now need to add in our distance data generated earlier and create a new column of total flows which **excludes flows that occur within areas**. 

*Note: we could keep the within-area (intra-area) flows, but they can cause problems so we will just exclude them for now.*

We first create a new total flows column which excludes intra-zone flow totals. We will set them to a very very small number to avoid making the intra-zonal distance become 0.

```{r echo=TRUE}
mdata$FlowNoIntra <- ifelse(mdata$Orig_code == mdata$Dest_code,0,mdata$Flow)
mdata$offset <- ifelse(mdata$Orig_code == mdata$Dest_code,0.0000000001,1)
```

We have ordered our spatial data earlier so that our zones are in their code order. We can now easily join these data together with our flow data as they are in the correct order.

```{r echo=TRUE}
mdata$dist <- distPair$value
```

and while we are here, rather than setting the intra-zonal distances to 0, we should set them to something small (most intrazonal moves won’t occur over 0 distance)

```{r echo=TRUE}
mdata$dist <- ifelse(mdata$dist == 0,5,mdata$dist)
```

We can now take a look at the migration dataset.

```{r echo=TRUE}
glimpse(mdata)
```

# Visualising with desire line

In this section, we learn how to prepare a desire line using **stplanr** package.

## Removing intra-zonal flows

We will not be plotting the intra-zonal flows. The code chunk below will be used to remove intra-zonal flows.

```{r echo=TRUE}
mdatasub <- mdata[mdata$Orig_code!=mdata$Dest_code,]
```

We first use *od2line()* of **stplanr** package to remove all but the origin, destination and flow columns.

```{r echo=TRUE}
mdatasub_skinny <- mdatasub[,c(2,4,5)]
travel_network <- od2line(flow = mdatasub_skinny, zones = Aus)
```

We will convert the flows to WGS84 projection next.

```{r echo=TRUE}
travel_networkwgs <- spTransform(travel_network,"+init=epsg:4326")
```

We then repeat the step for the `Aus` layer.

```{r echo=TRUE}
AusWGS <- spTransform(Aus,"+init=epsg:4326" )
```

Lastly, we will set the line widths to some sensible value according to the flow.

```{r echo=TRUE}
w <- mdatasub_skinny$Flow / max(mdatasub_skinny$Flow) * 10
```

We are now ready to plot the desire line map.

```{r echo=TRUE}
plot(travel_networkwgs, lwd = w)
plot(AusWGS, add=T)
```

# Building Spatial Interaction Models

We will now learn how to use R Stat function to calibrate the Spatial Interaction Models. Instead of using lm(), the **glm()** function will be used. This is because **glm()** allow us to calibrate the model using generalised linear regression methods.

## Unconstrained Spatial Interaction Model

We will calibrate an **unconstrained spatial interaction model** using *glm()*. The explanatory variables are **origin population** (i.e. vi1_origpop), **destination median income** (i.e. wj3_destmedinc) and **distance between origin and destination in km** (i.e. dist).

```{r echo=TRUE}
uncosim <- glm(Flow ~ log(vi1_origpop)+log(wj3_destmedinc)+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)

summary(uncosim)
```

### Fitting the model

To assess the performance of the model, we will use *fitted()* of **R** to compute the fitted values.

```{r echo=TRUE}
mdatasub$fitted <- fitted(uncosim)
```

### The more difficult way (optional)

Another way to calculate the estimates is to plug all of the parameters back into the Equation 6 model.

We first assign the parameter values from the model to the appropriate variables.

```{r echo=TRUE}
k <- uncosim$coefficients[1]
mu <- uncosim$coefficients[2]
alpha <- uncosim$coefficients[3]
beta <- -uncosim$coefficients[4]
```

We will next plug everything back into the Equation 6 model. 

Note: be careful with the positive and negative signing of the parameters as the beta parameter may not have been saved as negative. We will therefore need to force it to be negative.

```{r echo=TRUE}
mdatasub$unconstrainedEst2 <- exp(k+(mu*log(mdatasub$vi1_origpop))+(alpha*log(mdatasub$wj3_destmedinc))-(beta*log(mdatasub$dist)))
```

which is exactly the same as this

```{r echo=TRUE}
mdatasub$unconstrainedEst2 <- (exp(k)*exp(mu*log(mdatasub$vi1_origpop))*exp(alpha*log(mdatasub$wj3_destmedinc))*exp(-beta*log(mdatasub$dist)))
```

### Saving the fitted values

We will now run the model and save all of the new flow estimates in a new column in the data frame.

```{r echo=TRUE}
mdatasub$unconstrainedEst2 <- round(mdatasub$unconstrainedEst2,0)
sum(mdatasub$unconstrainedEst2)
```

We will next turn the output into a little matrix using *dcast()* of **maditr** package.

```{r echo=TRUE}
mdatasubmat2 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "unconstrainedEst2", margins=c("Orig_code", "Dest_code"))
mdatasubmat2
```

and compare with the original matrix.

```{r echo=TRUE}
mdatasubmat <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "Flow", margins=c("Orig_code", "Dest_code"))
mdatasubmat
```

We can also visualise the actual flow and estimated flow using a scatter plot.

```{r echo=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `unconstrainedEst2`))+
  geom_point(color="black", fill="light blue")
```

### Assessing the model performance

To provide a more formal assessment of the model, **Goodness-of-Fit** statistics will be used. *postReSample()* of **caret** package is used to compute three **Goodness-of-Fit** statistics.

```{r echo=TRUE}
postResample(mdatasub$Flow,mdatasub$unconstrainedEst2)
```

We can observe that the **R-squared value** is **0.32**, which is relatively low. It seems that the unconstrained model **failed to fit the empirical data well**.

## Origin Constrained Spatial Interaction Model

In this section, we will calibrate an **origin-constrained** SIM (the "-1" indicates no intercept in the regression model) using *glm()*.

```{r echo=TRUE}
origSim <- glm(Flow ~ Orig_code+log(wj3_destmedinc)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)

summary(origSim)
```

We can examine how the constraints hold for destinations this time.

Firstly, we will fit the model and roundup the estimated values.

```{r echo=TRUE}
mdatasub$origSimFitted <- round(fitted(origSim),0)
```

We will next use the step learnt earlier in the previous section to create a pivot table, to turn the paired list into a matrix.

```{r echo=TRUE}
mdatasubmat3 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "origSimFitted", margins=c("Orig_code", "Dest_code"))

mdatasubmat3
```

We can then compare with the original observed data.

```{r echo=TRUE}
mdatasubmat
```

We will next display the actual flow and estimated flow using a scatter plot.

```{r echo=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `origSimFitted`))+
  geom_point(color="black", fill="light blue")
```

Lastly, we compare the fitted values and the actual values by computing **Goodness-of-fit** statistics.

```{r echo=TRUE}
postResample(mdatasub$Flow,mdatasub$origSimFitted)
```
We can observe that the **R-squared** improved considerably from 0.32 in the unconstrained model to **0.43** in this **origin constrained model**.

## Destination Constrained Spatial Interaction Model

In this section, we will calibrate a **destination constrained** SIM (the "-1" indicates no intercept in the regression model) using *glm()*.

```{r echo=TRUE}
destSim <- glm(Flow ~ Dest_code+log(vi1_origpop)+log(dist)-1, na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)

summary(destSim)
```

We can examine how the constraints hold for destinations this time. 

Firstly, we will fit the model and roundup the estimated values.

```{r echo=TRUE}
mdatasub$destSimFitted <- round(fitted(destSim),0)
```

We will next use the step learnt earlier in the previous section to create a pivot table, to turn the paired list into a matrix.

```{r echo=TRUE}
mdatasubmat6 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "destSimFitted", margins=c("Orig_code", "Dest_code"))

mdatasubmat6
```

Similar to the previous section, we can then compare with the original observed data.

```{r echo=TRUE}
mdatasubmat
```

We will next display the actual flow and estimated flow using a scatter plot.

```{r echo=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `destSimFitted`))+
  geom_point(color="black", fill="light blue")
```

Finally, we can test the **Goodness-of-Fit**.

```{r echo=TRUE}
postResample(mdatasub$Flow,mdatasub$destSimFitted)
```

We can observe that the **R-squared** improved further from 0.32 in the unconstrained model to **0.65** in this **destination constrained model**.

## Doubly Constrained Spatial Interaction Model

In this section, we will calibrate a **Doubly Constrained** Spatial Interaction Model using *glm()*.

```{r echo=TRUE}
doubSim <- glm(Flow ~ Orig_code+Dest_code+log(dist), na.action = na.exclude, family = poisson(link = "log"), data = mdatasub)

summary(doubSim)
```

We can examine how the constraints hold for destinations this time. 

Firstly, we will fit the model and roundup the estimated values by using the code chunk below.

```{r echo=TRUE}
mdatasub$doubsimFitted <- round(fitted(doubSim),0)
```

Next, we will used the step learnt earlier in the previous section to create a pivot table, to turn the paired list into a matrix.

```{r echo=TRUE}
mdatasubmat7 <- dcast(mdatasub, Orig_code ~ Dest_code, sum, value.var = "doubsimFitted", margins=c("Orig_code", "Dest_code"))

mdatasubmat7
```

Similar to the previous section, we can then compare with the original observed data.

```{r echo=TRUE}
mdatasubmat
```

Next, we can display the actual flow and estimated flow using a scatter plot.
  
```{r echo=TRUE}
ggplot(data=mdatasub, 
       aes(y = `Flow`, 
           x = `doubsimFitted`))+
  geom_point(color="black", fill="light blue")
```

The scatter plot above reveals that the fitted values are **highly correlated** with the actual flow values. This show the Doubly Constrained Spatial Interaction Model is the **best fit model** among the four spatial interaction models.

To provide a quantitative assessment of the model, we can compute the **Goodness-of-fit** statistics.

```{r echo=TRUE}
postResample(mdatasub$Flow,mdatasub$doubsimFitted)
```

The **Goodness-of-fit** statistics reveal that the Doubly Constrained Spatial Interaction Model is the **best** model because it produces the **best R-squared statistic** and **smallest RMSE**.