---
title: "Hands-on Exercise 11"
description: |
  In this hands-on exercise, we learn how to model geographical accessibility using R’s geospatial analysis packages.
author:
  - name: Genice Goh
    url: {}
date: 10-30-2021
output:
  distill::distill_article:
    self_contained: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading Packages

The following packages will be used for this analysis:

- **sf**: used for spatial data handling
- **spatialAcc**: used for modelling geographical accessibility
- **tidyverse** (**readr**, **dplyr**): used for wrangling attribute data 
- **tmap**: used to prepare cartographic quality choropleth maps
- **ggplot2**: used for plotting statistical graphs
- **ggstatsplot**: used for statistical analysis

```{r echo=TRUE}
packages = c('tmap', 'SpatialAcc', 'sf', 'ggstatsplot', 'ggplot2', 'tidyverse')

for(p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p, character.only = T)
}
```

# Geospatial Data

## Importing Geospatial Data

*st_read()* of **sf** package is used to import the geospatial data, which is in **shapefile** format.

```{r echo=TRUE}
mpsz <- st_read(dsn = "data/geospatial", layer = "MP14_SUBZONE_NO_SEA_PL")
hexagons <- st_read(dsn = "data/geospatial", layer = "hexagons") 
eldercare <- st_read(dsn = "data/geospatial", layer = "ELDERCARE") 
```

The geometry type of the `mpsz` **sf** object is **multi-polygon**. It is also important to note that the `mpsz` **sf** object does not have EPSG information.

## Updating CRS Information

*st_transform()* of **sf** package is used to update the newly imported geospatial data with the correct ESPG code (i.e. 3414)

```{r echo=TRUE}
mpsz <- st_transform(mpsz, 3414)
eldercare <- st_transform(eldercare, 3414)
hexagons <- st_transform(hexagons, 3414)
```

After transforming the projection metadata, you can verify the projection of the newly transformed geospatial data using *st_crs()* of **sf** package.

```{r echo=TRUE}
st_crs(mpsz)
st_crs(eldercare)
st_crs(hexagons)
```

Notice that the EPSG code is indicated as 3414 now!

## Cleaning and Updating Attribute Fields

There are many redundant fields in the data tables of both `eldercare` and `hexagons`. We will be excluding them. At the same time, new fields called **demand** and **capacity** will be added into the data table of `hexagons` and `eldercare` **sf** data frame respectively. Both fields are derived using *mutate()* of **dplyr** package.

```{r echo=TRUE}
eldercare <- eldercare %>%
  select(fid, ADDRESSPOS) %>%
  mutate(capacity = 100)
```

```{r echo=TRUE}
hexagons <- hexagons %>%
  select(fid) %>%
  mutate(demand = 100)
```

Notice that for the purpose of this hands-on exercise, a constant value of 100 is used. In practice, **actual** demand of the hexagon and capacity of the eldercare centre should be used.

# Aspatial Data

## Importing Distance Matrix

*read_csv()* of **readr** package is used to import **OD_Matrix.csv**. The imported object is a tibble data frame called `ODMatrix`.

```{r echo=TRUE}
ODMatrix <- read_csv("data/aspatial/OD_Matrix.csv", skip = 0)
```

## Tidying Distance Matrix

The imported ODMatrix organised the distance matrix columnwise.

On the other hands, most of the modelling packages in R is expecting a matrix where the **rows** represent the **origins** (i.e. from field) and the **columns** represent the **destinations** (i.e. to field)

*spread()* of **tidyr** package is used to transform the O-D matrix from a thin format into a fat format.

Note: *pivot_wider()* is newly introduced and should be used instead of *spread()*.

```{r echo=TRUE}
distmat <- ODMatrix %>%
  select(origin_id, destination_id, total_cost) %>%
  spread(destination_id, total_cost)%>%
  select(c(-c('origin_id')))
```

Currently, the distance is measured in **metre** because SVY21 projected coordinate system is used. We will need to convert the unit of measurement from **metre to kilometre**.

```{r echo=TRUE}
distmat_km<-as.matrix(distmat/1000)
```

# Modelling and Visualising Accessibility using Hansen Method

## Computing Hansen’s Accessibility

We are now ready to compute Hansen’s accessibility using *ac()* of **SpatialAcc** package. *data.frame()* is then used to save the output in a data frame called `acc_Handsen`.

```{r echo=TRUE}
acc_Hansen <- data.frame(ac(hexcen$demand, eldercare$capacity, distmat_km, 
                            d0 = 50, power = 2, family = "Hansen"))
```

The default field name is very messy, we will rename it to **accHansen**.

```{r echo=TRUE}
colnames(acc_Hansen) <- "accHansen"
```

Notice that the field name is much more tidier now.

Next, we will convert the data table into **tibble** format.

```{r echo=TRUE}
acc_Hansen <- as_tibble(acc_Hansen)
```

Lastly, *bind_cols()* of **dplyr** package will be used to join the `acc_Hansen` **tibble** data frame with the `hexagons` **sf** data frame. The output is called `hexagon_Hansen`.

```{r echo=TRUE}
hexagon_Hansen <- bind_cols(hexagons, acc_Hansen)
```

Notice that `hexagon_Hansen` is a **sf** data frame and not a typical **tibble** data frame.

## Visualising Hansen’s Accessibility

### Extracting Map Extend

Firstly, we will extract the extent of `hexagons` **sf** data frame using *st_bbox()* of **sf** package.

```{r echo=TRUE}
mapex <- st_bbox(hexagons) #view extent
```

We will now use a collection of mapping functions of **tmap** package to create a high cartographic quality accessibility to eldercare centre in Singapore.

```{r echo=TRUE, fig.width=12, fig.height=8}
tm_shape(hexagon_Hansen,
         bbox = mapex) + 
  tm_fill(col = "accHansen",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: Hansen method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical Graphic Visualisation

In this section, we are going to compare the distribution of Hansen’s accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into `haxegon_Hansen` **sf** data frame.

```{r echo=TRUE}
hexagon_Hansen <- st_join(hexagon_Hansen, mpsz, 
                          join = st_intersects)
```

Next, *ggplot()* will be used to plot the distribution using the boxplot graphical method.

```{r echo=TRUE, fig.width=8}
ggplot(data=hexagon_Hansen, 
       aes(y = log(accHansen), 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# Modelling and Visualising Accessibility using KD2SFCA Method

## Computing KD2SFCA’s Accessibility

We will calculate Hansen’s accessibility using *ac()* of **SpatialAcc** package and *data.frame()* is used to save the output in a data frame called `acc_KD2SFCA`. Notice that **KD2SFCA** is used for **family argument**.

Next, we will convert the data table into **tibble** format.

Lastly, *bind_cols()* of **dplyr** package will be used to join the `acc_KD2SFCA` **tibble** data frame with the `hexagons` **sf** data frame. The output is called `hexagon_KD2SFCA`.

```{r echo=TRUE}
acc_KD2SFCA <- data.frame(ac(hexagons$demand, eldercare$capacity, distmat_km, 
                            d0 = 50, power = 2, family = "KD2SFCA"))

colnames(acc_KD2SFCA) <- "accKD2SFCA"
acc_KD2SFCA <- as_tibble(acc_KD2SFCA)
hexagon_KD2SFCA <- bind_cols(hexagons, acc_KD2SFCA)
```

## Visualising KD2SFCA’s Accessibility

We will now use a collection of mapping functions of **tmap** package to create a high cartographic quality accessibility to eldercare centre in Singapore. Notice that `mapex` is reused for the **bbox** argument.

```{r echo=TRUE, fig.width=10, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_KD2SFCA,
         bbox = mapex) + 
  tm_fill(col = "accKD2SFCA",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: KD2SFCA method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 6),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical Graphic Visualisation

Now, we are going to compare the distribution of KD2CFA accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into `hexagon_KD2SFCA` **sf** data frame.

```{r echo=TRUE}
hexagon_KD2SFCA <- st_join(hexagon_KD2SFCA, mpsz, 
                          join = st_intersects)
```

Next, *ggplot()* will be used to plot the distribution using the boxplot graphical method.

```{r echo=TRUE, fig.width=8}
ggplot(data=hexagon_KD2SFCA, 
       aes(y = accKD2SFCA, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```

# Modelling and Visualising Accessibility using Spatial Accessibility Measure (SAM) Method

## Computing SAM Accessibility

We will calculate Hansen’s accessibility using *ac()* of **SpatialAcc** package and *data.frame()* is used to save the output in a data frame called `acc_SAM`. Notice that **SAM** is used for **family argument**.

Next, we will convert the data table into **tibble** format.

Lastly, *bind_cols()* of **dplyr** package will be used to join the `acc_SAM` **tibble** data frame with the `hexagons` **sf** data frame. The output is called `hexagon_SAM`.

```{r echo=TRUE}
acc_SAM <- data.frame(ac(hexagons$demand, eldercare$capacity, distmat_km, 
                         d0 = 50, power = 2, family = "SAM"))

colnames(acc_SAM) <- "accSAM"
acc_SAM <- as_tibble(acc_SAM)
hexagon_SAM <- bind_cols(hexagons, acc_SAM)
```

## Visualising SAM’s Accessibility

We will now use a collection of mapping functions of **tmap** package to create a high cartographic quality accessibility to eldercare centre in Singapore. Notice that `mapex` is reused for **bbox** argument.

```{r echo=TRUE, fig.width=10, fig.height=8}
tmap_mode("plot")
tm_shape(hexagon_SAM,
         bbox = mapex) + 
  tm_fill(col = "accSAM",
          n = 10,
          style = "quantile",
          border.col = "black",
          border.lwd = 1) +
tm_shape(eldercare) +
  tm_symbols(size = 0.1) +
  tm_layout(main.title = "Accessibility to eldercare: SAM method",
            main.title.position = "center",
            main.title.size = 2,
            legend.outside = FALSE,
            legend.height = 0.45, 
            legend.width = 3.0,
            legend.format = list(digits = 3),
            legend.position = c("right", "top"),
            frame = TRUE) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar(width = 0.15) +
  tm_grid(lwd = 0.1, alpha = 0.5)
```

## Statistical Graphic Visualisation

Now, we are going to compare the distribution of SAM accessibility values by URA Planning Region.

Firstly, we need to add the planning region field into `hexagon_SAM` **sf** data frame.

```{r echo=TRUE}
hexagon_SAM <- st_join(hexagon_SAM, mpsz, 
                       join = st_intersects)
```

Next, *ggplot()* will be used to plot the distribution using the boxplot graphical method.

```{r echo=TRUE, fig.width=8}
ggplot(data=hexagon_SAM, 
       aes(y = accSAM, 
           x= REGION_N)) +
  geom_boxplot() +
  geom_point(stat="summary", 
             fun.y="mean", 
             colour ="red", 
             size=2)
```