---
title: "In-class Exercise 5"
description: |
  This in-class exercise is a revision of what is covered in Hands-on Exercise 4, which revolves around spatial point patterns analysis.
author:
  - name: Genice Goh
    url: {}
date: 09-13-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Installing and Loading Packages

- maptools: creating spatial* objects
- sf: importing geospatial data
- raster: creating raster object
- spatstat: performing SPPA
- tmap: plotting thematic maps
- tidyverse: data handling
- plotly: creating interactive plots
- ggthemes: provides more advanced themes for ggplot

```{r echo=TRUE}
packages = c('sf', 'spatstat', 'raster', 'maptools', 'tmap', 'tidyverse', 'plotly', 'ggthemes')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## Importing Geospatial Data

- *st_read()* of **sf* package is for importing geospatial data; not appropriate for importing aspatial data.
- The output object is in tibble sf object class.

```{r echo=TRUE}
mpsz_sf <- st_read(dsn="data/shapefile",
                   layer="MP14_SUBZONE_WEB_PL")
```

- Projection is in SVY21.

## Importing Aspatial Data

*read_rds()* of readr package is used instead of *readRDS()* of base R is used. This is because the output of *read_rds()* is in tibble object.

```{r echo=TRUE}
childcare <- read_rds("data/rds/childcare.rds")
CHAS <- read_rds("data/rds/CHAS.rds")
```

Note: There are some data issues in childcare dataframe because `Lat` and `Lng` should be in numeric data type. The coordinate fields seem to be in decimal degrees and it is assumed that WGS84 referencing system is used.

## Convert Aspatial Data into sf objects

```{r echo=TRUE}
CHAS_sf <- st_as_sf(CHAS, 
                    coords=c("X_COORDINATE", "Y_COORDINATE"),
                    crs=3414)
```

Note: *st_as_sf()* accepts coordinates in character data type

```{r echo=TRUE}
childcare_sf <- st_as_sf(childcare, 
                    coords=c("Lng", "Lat"),
                    crs=4326) %>%
  st_transform(crs=3414)
```

- No significant difference if the columns are not numeric during conversion. But we should convert the columns to numeric type for best practice. 

## Plotting for Reviewing

- alpha: set transparency of dots
- color: set colour of dots
- size: set size of dots

```{r echo=TRUE}
tmap_mode('view')
tm_shape(childcare_sf) + 
  tm_dots(alpha=0.4,
          col="blue",
          size= 0.05) +
tm_shape(CHAS_sf) + 
  tm_dots(alpha=0.4,
          col="red",
          size= 0.05)
```

## Geospatial Data Wrangling

```{r}
tmap_mode('plot')
```

### Converting from sf to Spatial* dataframe

 Using *as_Spatial()* of sf package

- coordinates: geometries object
- bbox: extent of geometries
- proj4string: defines projection

```{r echo=TRUE}
childcare <- as_Spatial(childcare_sf)
CHAS <- as_Spatial(CHAS_sf)
mpsz <- as_Spatial(mpsz_sf)
```

### Converting Spatial* dataframe into Spatial* objects

Using *as.SpatialPoint()* / *as.SpatialPolygon()* of maptools package

- Input (e.g. childcare) needs to be a Spatial* dataframe
- Output object is **not a dataframe** anymore since it does not contain data object

```{r echo=TRUE}
childcare_sp <- as(childcare, "SpatialPoints")
CHAS_sp <- as(CHAS, "SpatialPoints")
mpsz_sp <- as(mpsz, "SpatialPolygons")
```

### Converting Spatial* objects into ppp objects

Using *as.ppp()* function of maptools package 

- Drops away projection information

```{r echo=TRUE}
childcare_ppp <- as(childcare_sp, "ppp")
CHAS_ppp <- as(CHAS_sp, "ppp")
```

### Remove Duplicate Points using Jitter

```{r echo=TRUE}
childcare_ppp_jit <- rjitter(childcare_ppp,
                             retry=TRUE,
                             nsim=1,
                             drop=TRUE)

any(duplicated(childcare_ppp_jit))
```

```{r echo=TRUE}
CHAS_ppp_jit <- rjitter(CHAS_ppp,
                             retry=TRUE,
                             nsim=1,
                             drop=TRUE)

any(duplicated(CHAS_ppp_jit))
```

- Cannot use tmap() to plot ppp objects for visualisation. Need to convert the ppp object to sf object before it is possible to plot using tmap(). Will have to resort to using plot()

### Extract Punggol Planning Area

- @: to access data object 
- $: to access columns in data object

```{r echo=TRUE}
pg <- mpsz[mpsz@data$PLN_AREA_N=="PUNGGOL",]
```

### Convert SpatialPolygonsDataFrame into Spatial Polygons object

```{r echo=TRUE}
pg_sp <- as(pg, "SpatialPolygons")
```

### Convert SpatialPolygons into owin object

```{r echo=TRUE}
pg_owin <- as(pg_sp, "owin")
```

### Extract Spatial Points within Owin

- Points will now within the Punggol sub-district

```{r echo=TRUE}
childcare_pg <- childcare_ppp_jit[pg_owin]
CHAS_pg <- CHAS_ppp_jit[pg_owin]
```

```{r echo=TRUE}
plot(childcare_pg)
```

## L-function

```{r echo=TRUE}
L_childcare <- envelope(childcare_pg,
                        Lest,
                        nsim=99,
                        rank=1,
                        global=TRUE)
```

```{r echo=TRUE}
L_CHAS <- envelope(CHAS_pg,
                        Lest,
                        nsim=99,
                        rank=1,
                        global=TRUE)
```

```{r echo=TRUE}
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_childcare)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
}
```

```{r echo=TRUE} 
title <- "Pairwise Distance: L function"

Lcsr_df <- as.data.frame(L_CHAS)

colour=c("#0D657D","#ee770d","#D3D3D3")
csr_plot <- ggplot(Lcsr_df, aes(r, obs-r))+
  # plot observed value
  geom_line(colour=c("#4d4d4d"))+
  geom_line(aes(r,theo-r), colour="red", linetype = "dashed")+
  # plot simulation envelopes
  geom_ribbon(aes(ymin=lo-r,ymax=hi-r),alpha=0.1, colour=c("#91bfdb")) +
  xlab("Distance r (m)") +
  ylab("L(r)-r") +
  geom_rug(data=Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,], sides="b", colour=colour[1])  +
  geom_rug(data=Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,], sides="b", colour=colour[2]) +
  geom_rug(data=Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,], sides="b", color=colour[3]) +
  theme_tufte()+
  ggtitle(title)

text1<-"Significant clustering"
text2<-"Significant segregation"
text3<-"Not significant clustering/segregation"

# the below conditional statement is required to ensure that the labels (text1/2/3) are assigned to the correct traces
if (nrow(Lcsr_df[Lcsr_df$obs > Lcsr_df$hi,])==0){ 
  if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text3, traces = 4) %>%
      rangeslider() 
  }else if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){ 
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      rangeslider() 
  }else {
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text2, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider() 
  }
} else if (nrow(Lcsr_df[Lcsr_df$obs < Lcsr_df$lo,])==0){
  if (nrow(Lcsr_df[Lcsr_df$obs >= Lcsr_df$lo & Lcsr_df$obs <= Lcsr_df$hi,])==0){
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      rangeslider() 
  } else{
    ggplotly(csr_plot, dynamicTicks=T) %>%
      style(text = text1, traces = 4) %>%
      style(text = text3, traces = 5) %>%
      rangeslider()
  }
} else{
  ggplotly(csr_plot, dynamicTicks=T) %>%
    style(text = text1, traces = 4) %>%
    style(text = text2, traces = 5) %>%
    style(text = text3, traces = 6) %>%
    rangeslider()
}
```