---
title: "Take-home Exercise 1"
description: |
  A short description of the post.
author:
  - name: Genice Goh
    url: {}
date: 09-10-2021
output:
  distill::distill_article:
    self_contained: false
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## 1.0 Introduction
This take-home exercise aims to reveal the spatial-temporal patterns of monthly cumulative confirmed COVID-19 rates and death rates in Jakarta DKI at the sub-district level.

## 2.0 The Data
For this assignment, the following data is used:

- Daily COVID-19 data at both sub-district and distict level. It is extracted from [Open Data Covid-19 Provinsi DKI Jakarta](https://riwayat-file-covid-19-dki-jakarta-jakartagis.hub.arcgis.com/).
- Geospatial data of Jakarta DKI in the form of a Shapefile (SHP). It is extracted from [PODES 2019](https://www.indonesia-geospasial.com/2020/04/download-shapefile-shp-batas-desa.html).

## 3.0 Instaling and Loading Packages

In this take-home exercise, the packages sf, tidyverse, tmap and readxl are used. 

```{r echo=TRUE, eval=TRUE}
packages = c('sf', 'tidyverse', 'tmap', 'readxl')
for (p in packages){
  if(!require(p, character.only = T)){
    install.packages(p)
  }
  library(p,character.only = T)
}
```

## 4.0 Data Preparation and Wrangling
### 4.1 Geospatial Data

The geospatial data is imported with the use of *st_read()*.

```{r echo=TRUE, eval=TRUE}
DKI <- st_read(dsn="data/geospatial",
               layer="BATAS_DESA_DESEMBER_2019_DUKCAPIL_DKI_JAKARTA")

glimpse(DKI)
```

### 4.1.1 Checking for NA values

Since the value TRUE is returned, it implies that the dataset contains NA values.

```{r echo=TRUE, eval=TRUE}
sum(is.na(DKI)) > 0
```

The function *drop_na* is used to drop the NA values in the dataset.

```{r echo=TRUE, eval=TRUE}
DKI <- drop_na(DKI)
```

### 4.1.2 CRS Check and Reprojection

The *st_crs()* function is used to check the coordinate reference system of the dataset.

```{r echo=TRUE, eval=TRUE}
st_crs(DKI)
```

Since the dataset is in WGS84 projection system, there is a need to transform it to the national projected coordinates system of Indonesia (DGN95/Indonesia TM-3 zone 54.1), which has an EPSG code of 23845.

```{r echo=TRUE, eval=TRUE}
DKI23845 <- st_transform(DKI, crs = 23845)
st_crs(DKI23845)
```

### 4.1.3 Exclude all outer islands

The *tmap()* function is used to plot the geospatial data for easy visualisation.

```{r echo=TRUE, eval=TRUE}
tmap_mode("view")
tm_shape(DKI23845) + 
  tm_polygons()
```

While looking at the data points of the outer islands using the interactive map, it can be observed that they share the same value "KEPULAUAN SERIBU" for the field "KAB_KOTA". This information can therefore be used to exclude the outer islands from the dataframe.

```{r echo=TRUE, eval=TRUE}
tmap_mode('plot')

DKI23845 <- DKI23845 %>% 
  filter(`KAB_KOTA` != "KEPULAUAN SERIBU")

plot(DKI23845)
```

### 4.1.4 Select the first 9 fields
```{r echo=TRUE, eval=TRUE}
DKI23845 <- DKI23845 %>% select(0:9)

DKI23845
```

### 4.2 Aspatial Data

The aspatial data is imported using the *read_excel* function. Since data is only required at the sub-district level, data will only be extracted from the "data" sheet.

```{r echo=TRUE, eval=FALSE}
mar20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (31 Maret 2020 Pukul 08.00).xlsx", sheet="data")

apr20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (30 April 2020 Pukul 09.00).xlsx", sheet="data") 

may20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (31 MEI 2020 Pukul 09.00).xlsx", sheet="data") 

jun20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (30 Juni 2020 Pukul 09.00).xlsx", sheet="data") 

jul20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (31 Juli 2020 Pukul 09.00).xlsx", sheet="data") 

aug20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (31 Agustus 2020 Pukul 10.00).xlsx", sheet="data") 

sep20 <- read_excel("data/aspatial/Standar_Kelurahan_Data_Corona_30_September_2020_Pukul_10_00.xlsx", sheet="data") 

oct20 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (31 Oktober 2020 Pukul 10.00).xlsx", sheet="data") 

nov20 <- read_excel("data/aspatial/Standar_Kelurahan_Data_Corona_30_November_2020_Pukul_10_00.xlsx", sheet="data") 

dec20 <- read_excel("data/aspatial/Standar_Kelurahan_Data_Corona_26_Desember_2020_Pukul_10_00.xlsx", sheet="data") 

jan21 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (30 Januari 2021 Pukul 10.00).xlsx", sheet="data") 

feb21 <- read_excel("data/aspatial/Standar_Kelurahan_Data_Corona_27_Februari_2021_Pukul_10_00.xlsx", sheet="data")

mar21 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (27 Maret 2021 Pukul 10.00).xlsx", sheet="data") 

apr21 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (24 April 2021 Pukul 10.00).xlsx", sheet="data")

may21 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (29 Mei 2021 Pukul 10.00).xlsx", sheet="data") 

jun21 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (26 Juni 2021 Pukul 10.00).xlsx", sheet="data") 

jul21 <- read_excel("data/aspatial/Standar Kelurahan Data Corona (31 Juli 2021 Pukul 10.00).xlsx", sheet="data")
```


### 4.2.1 Extracting Required Columns

In the case of duplicated columns, values from the 2nd column are retained for consistency. There is also a need to create a new column called Month, with values following the respective months.

```{r echo=TRUE, eval=FALSE}
mar20_filtered <- mar20 %>% 
  select("ID_KEL"="ID_KEL...2", "Nama_provinsi", "nama_kota", 
         "nama_kecamatan", "nama_kelurahan", "POSITIF", "Meninggal") %>%
  mutate("Month"= "Mar20")

apr20_filtered <- apr20 %>% 
  select("ID_KEL"="ID_KEL...2", "Nama_provinsi", "nama_kota", 
         "nama_kecamatan", "nama_kelurahan", "POSITIF", "Meninggal") %>%
  mutate("Month"= "Apr20")

may20_filtered <- may20 %>% 
  select("ID_KEL"="ID_KEL...2", "Nama_provinsi", "nama_kota", 
         "nama_kecamatan", "nama_kelurahan", "POSITIF", "Meninggal") %>%
  mutate("Month"= "May20")

jun20_filtered <- jun20 %>% 
  select("ID_KEL"="ID_KEL...2", "Nama_provinsi", "nama_kota", 
         "nama_kecamatan", "nama_kelurahan", "POSITIF", "Meninggal") %>%
  mutate("Month"= "Jun20")

jul20_filtered <- jul20 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...26") %>%
  mutate("Month"= "Jul20")

aug20_filtered <- aug20 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...28") %>%
  mutate("Month"= "Aug20")

sep20_filtered <- sep20 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...29") %>%
  mutate("Month"= "Sep20")

oct20_filtered <- oct20 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...30") %>%
  mutate("Month"= "Oct20")

nov20_filtered <- nov20 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...30") %>%
  mutate("Month"= "Nov20")

dec20_filtered <- dec20 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...30") %>%
  mutate("Month"= "Dec20")

jan21_filtered <- jan21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "Jan21")

feb21_filtered <- feb21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan", 
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "Feb21")

mar21_filtered <- mar21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan",
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "Mar21")

apr21_filtered <- apr21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan",
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "Apr21")

may21_filtered <- may21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan",
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "May21")

jun21_filtered <- jun21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan",
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "Jun21")

jul21_filtered <- jul21 %>% 
  select("ID_KEL", "Nama_provinsi", "nama_kota", "nama_kecamatan",
         "nama_kelurahan", "POSITIF", "Meninggal"="Meninggal...31") %>%
  mutate("Month"= "Jul21")
```

### 4.2.2 Integrating monthly data
```{r echo=TRUE, eval=FALSE}
covidDKI <- bind_rows(mar20_filtered, apr20_filtered, may20_filtered,
                      jun20_filtered, jul20_filtered, aug20_filtered,
                      sep20_filtered, oct20_filtered, nov20_filtered,
                      dec20_filtered, jan21_filtered, feb21_filtered,
                      mar21_filtered, apr21_filtered, may21_filtered,
                      jun21_filtered, jul21_filtered)

covidDKI
```

### 4.2.3 Checking for NA values

Since the value TRUE is returned, it implies that the dataset contains NA values.

```{r echo=TRUE, eval=FALSE}
sum(is.na(covidDKI)) > 0
```


The function *drop_na* is used to drop the NA values in the dataset.


```{r echo=TRUE, eval=FALSE}
covidDKI <- drop_na(covidDKI)
```

### 4.2.4 Data Cleaning

While looking through the cleaned dataset, it can be observed there are rows where the ID_KEL column has non-integer values. These observations are therefore removed.

```{r echo=TRUE, eval=FALSE}
covidDKI <- covidDKI %>%
  filter(`ID_KEL` != "LUAR DKI JAKARTA") %>% 
  filter(`ID_KEL` != "PROSES UPDATE DATA") %>% 
  filter(`ID_KEL` != "BELUM DIKETAHUI")
```

Moreover, several data points in the column "nama_kelurahan" has values in both the short and long form. There is therefore a need to convert those values in the short form to the long form to aid later data wrangling. 

```{r echo=TRUE, eval=FALSE}
covidDKI$nama_kelurahan <- gsub("P\\.", "PULAU", covidDKI$nama_kelurahan)
```

### 4.2.5 Save the aspatial data as RDS 
```{r echo=TRUE, eval=FALSE}
covidDKI_rds <- write_rds(covidDKI, "data/rds/covidDKI.rds")
```

The RDS aspatial data is now used to proceed with the later analysis.

```{r echo=TRUE, eval=TRUE}
covidDKI <- read_rds("data/rds/covidDKI.rds")
```

### 4.2.6 Prepare cumulative monthly data in every column

Since it is required later to plot the cumulative confirmed cases rates and death rates by month, the dataset is pivoted to a wide version. 

```{r echo=TRUE, eval=TRUE}
covidDKI <- covidDKI %>%
  group_by(ID_KEL, Nama_provinsi, nama_kota, nama_kecamatan, 
           nama_kelurahan, Month) %>%
  summarise(`MENINGGAL` = sum(`Meninggal`), `POSITIF` = sum(`POSITIF`)) %>%
  ungroup() %>%
  pivot_wider(names_from=Month, values_from=c(POSITIF, MENINGGAL))
```

### 4.3 Combine aspatial and geospatial data

The function *right_join()* is used to combine the aspatial and geospatial data.

```{r echo=TRUE, eval=TRUE}
jakartaDKI <- right_join(covidDKI, DKI23845,
                         by= c("ID_KEL" = "KODE_DESA"))
```

### 4.3.1 Checking for NA values

Since the value returned is FALSE, there are no NA vlues in the dataset.

```{r echo=TRUE}
sum(is.na(jakartaDKI)) > 0
```

### 4.4 Calculate cumulative confirmed cases and death rates by month

There is a need to generate the total number of confirmed cases and deaths in each area using *rowSums()* and generate 2 new columns based on these values using *mutate()*.

```{r echo=TRUE, eval=TRUE}
jakartaDKI <- jakartaDKI %>%
  mutate(`TOTAL POSITIF` = rowSums(.[6:22])) %>%
  mutate(`TOTAL MENINGGAL` = rowSums(.[23:39]))
```

The cumulative confirmed cases and death rates (per 10000 population) by month is then calculated.

```{r echo=TRUE, eval=TRUE}
jakartaDKI_rates <- jakartaDKI 

jakartaDKI_rates[6:22] <- 
  (jakartaDKI_rates[6:22] / jakartaDKI_rates$JUMLAH_PEN) * 10000

jakartaDKI_rates[23:39] <-
  (jakartaDKI_rates[23:39] / jakartaDKI_rates$JUMLAH_PEN) * 10000
```

### 4.5 Calculate Relative Risk
```{r echo=TRUE, eval=TRUE} 
jakartaDKI_risk <- jakartaDKI %>%
  mutate(`RELATIVE_RISK` = (rowSums(.[23:39]) * 100) / 
           (rowSums(.[23:39]) * jakartaDKI$JUMLAH_PEN))
```

## 5.0 Thematic Mapping

Choropleth mapping techniques will be used to analyse the spatial-temporal patterns of monthly cumulative confirmed cases rates and death rates in Jakarta DKI at the sub-district level.

### 5.1 Converting datasets to sf format

The required datasets need to be converted to sf format using *st_as_sf()* so that we can create plots using tmap.

```{r echo=TRUE, eval=TRUE}
jakartaDKI <- st_as_sf(jakartaDKI)
jakartaDKI_rates <- st_as_sf(jakartaDKI_rates)
jakartaDKI_risk <- st_as_sf(jakartaDKI_risk)
```

### 5.2 Classification Method

In order to have a common basis of comparison among the choropleth maps, they need to use the same classification method. 

The equal intervals classification method with custom breaks will be utilised to plot the choropleth map.  

### 5.2.1 Selecting Custom Breaks

To specify the intervals of the break points, descriptive statistics of the cumulative confirmed cases rates and deaths rates are examined. 

```{r echo=TRUE, eval=TRUE}
summary(jakartaDKI_rates$POSITIF_Mar20)
```

```{r echo=TRUE, eval=TRUE}
summary(jakartaDKI_rates$POSITIF_Jul21)
```

- The cumulative confirmed cases rates range from 0 to 3808 across the sub-districts from Mar20 to Jul21.
- Break points can therefore be defined to be in intervals of 476, with a total number of 8 classes.

```{r echo=TRUE, eval=TRUE}
summary(jakartaDKI_rates$MENINGGAL_Mar20)
```

```{r echo=TRUE, eval=TRUE}
summary(jakartaDKI_rates$MENINGGAL_Jul21)
```

- The cumulative death rates range from 0 to 42 across the sub-districts from Mar20 to Jul21.
- Break points can therefore be defined to be in intervals of 6, with a total number of 7 classes.

For both cumulative confirmed cases rates and death rates, it should be kept in mind that there may be over-generalisation of map data due to the intervals of the break points. This is especially during the earlier months where the cumulative confirmed cases rates and death rates are both low, and important spatial trends may have been missed. 

As a result, the Jenks Classification Method will be utilised to plot the choropleth maps as well. This is because it allows for natural breaks, where break points are specified in a way that maximises within-group homogeneity and between-group heterogeneity. This can help us to capture potential critical spatial trends that we may have overlooked previously.  

### 5.3 Equal Intervals Classification Method

The *equal_positif* and *equal_meninggal* functions are created to perform choropleth mapping using the equal intervals classification method with custom breaks. 

As mentioned earlier, there are a total number of 8 classes defined, with an interval of 476 for the cumulative confirmed cases rates. 

```{r echo=TRUE, eval=TRUE}
equal_positif <- function(df, vnam) {
  tm_shape(jakartaDKI_rates)+
  tm_fill(vnam,
          breaks = c(0, 476, 952, 1428, 1904, 2380, 2856, 3332, 3808),
          palette="Blues") +
  tm_borders(lwd = 0.1,
             alpha = 0.3) + 
  tm_layout(panel.show = TRUE,
            panel.labels = vnam,
            panel.label.size = 0.5,
            legend.show=FALSE)
}
```

For the cumulative death rates, there are a total number of 7 classes defined, with an interval of 6. 

```{r echo=TRUE, eval=TRUE}
equal_meninggal <- function(df, vnam) {
  tm_shape(jakartaDKI_rates)+
  tm_fill(vnam,
          breaks = c(0, 6, 12, 18, 24, 30, 36, 42),
          palette="Blues") +
  tm_borders(lwd = 0.1,
             alpha = 0.3) + 
  tm_layout(panel.show = TRUE,
            panel.labels = vnam,
            panel.label.size = 0.5,
            legend.show=FALSE)
}
```

As seen in the code chunks below, the *equal_positif* and *equal_meninggal* functions can be easily applied to create choropleth maps for the cumulative confirmed cases rates and death rates without much duplicate code. The choropleth maps are then arranged together using *tmap_arrange()* to aid easier spotting of patterns over space and time. 

### 5.3.1 Cumulative Confirmed Cases Rates By Month
```{r echo=TRUE, eval=TRUE}
tmap_arrange(equal_positif(jakartaDKI_rates, "POSITIF_Mar20"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Apr20"),
             equal_positif(jakartaDKI_rates, "POSITIF_May20"),
             equal_positif(jakartaDKI_rates, "POSITIF_Jun20"),
             equal_positif(jakartaDKI_rates, "POSITIF_Jul20"),
             equal_positif(jakartaDKI_rates, "POSITIF_Aug20"),
             equal_positif(jakartaDKI_rates, "POSITIF_Sep20"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Oct20"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Nov20"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Dec20"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Jan21"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Feb21"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Mar21"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Apr21"), 
             equal_positif(jakartaDKI_rates, "POSITIF_May21"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Jun21"), 
             equal_positif(jakartaDKI_rates, "POSITIF_Jul21"))
```

- A darker shade of blue signifies higher cumulative confirmed cases rates.
- Not much spatial patterns can be observed from March 2020 to January 2021.
- It can be observed that only 1 sub-district in the central region has higher cumulative confirmed cases rates in February 2021.
- However, sub-districts in the surroundings begin to have higher cumulative confirmed cases rates across time as well, which can be seen from March 2021 to July 2021.
- In the month of July 2021, it can be seen that the sub-districts with high cumulative confirmed cases rates are concentrated in the central region, while other sub-districts are also observed to have higher cumulative confirmed cases rates as compared to March 2020.

### 5.3.2 Cumulative Death Rates By Month
```{r echo=TRUE, eval=TRUE}
tmap_arrange(equal_meninggal(jakartaDKI_rates, "MENINGGAL_Mar20"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Apr20"),
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_May20"),
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Jun20"),
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Jul20"),
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Aug20"),
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Sep20"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Oct20"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Nov20"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Dec20"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Jan21"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Feb21"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Mar21"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Apr21"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_May21"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Jun21"), 
             equal_meninggal(jakartaDKI_rates, "MENINGGAL_Jul21"))
```

- A darker shade of blue signifies higher cumulative death rates.
- Not much spatial patterns can be observed from March 2020 to January 2021.
- Similarly, there is 1 sub-district in the central region with higher cumulative death rates in February 2021. It can also be observed from March 2021 to July 2021 that the surrounding sub-districts begin to have higher cumulative death rates as well.
- The trend identified earlier is also observed here: in July 2021, it can be seen that the sub-districts with high cumulative death rates are concentrated in the central region, while other sub-districts are also observed to have higher cumulative death rates as compared to March 2020.

Similar to what we have predicted before, we cannot identify significant spatial patterns from March 2020 to January 2021, which may have been lost due to the break points defined. As a result, choropleth maps using the Jenks classification method will be created to give further insights.

### 5.4 Jenks Classification Method

The *jenks* function is created to perform choropleth mapping using the Jenks classification method. 

```{r echo=TRUE, eval=TRUE}
jenks <- function(df, vnam){
  tm_shape(df) +
     tm_fill(vnam,
             n = 5,
             style = "jenks",
             palette="Blues")  +
  tm_borders(lwd = 0.1,
             alpha = 0.3) +
  tm_layout(panel.show = TRUE,
            panel.labels = vnam,
            panel.label.size = 0.5,
            legend.show=FALSE)
}
```

As seen in the code chunks below, the *jenks* function can be easily applied to create choropleth maps for the cumulative confirmed cases rates and death rates without much duplicate code. The choropleth maps are then arranged together using *tmap_arrange()* to aid easier spotting of patterns over space and time.

### 5.4.1 Cumulative Confirmed Cases Rates By Month
```{r echo=TRUE, eval=TRUE}
tmap_arrange(jenks(jakartaDKI_rates, "POSITIF_Mar20"), 
             jenks(jakartaDKI_rates, "POSITIF_Apr20"),
             jenks(jakartaDKI_rates, "POSITIF_May20"),
             jenks(jakartaDKI_rates, "POSITIF_Jun20"),
             jenks(jakartaDKI_rates, "POSITIF_Jul20"),
             jenks(jakartaDKI_rates, "POSITIF_Aug20"),
             jenks(jakartaDKI_rates, "POSITIF_Sep20"), 
             jenks(jakartaDKI_rates, "POSITIF_Oct20"), 
             jenks(jakartaDKI_rates, "POSITIF_Nov20"), 
             jenks(jakartaDKI_rates, "POSITIF_Dec20"), 
             jenks(jakartaDKI_rates, "POSITIF_Jan21"), 
             jenks(jakartaDKI_rates, "POSITIF_Feb21"), 
             jenks(jakartaDKI_rates, "POSITIF_Mar21"), 
             jenks(jakartaDKI_rates, "POSITIF_Apr21"), 
             jenks(jakartaDKI_rates, "POSITIF_May21"), 
             jenks(jakartaDKI_rates, "POSITIF_Jun21"), 
             jenks(jakartaDKI_rates, "POSITIF_Jul21"))
```

- A darker shade of blue signifies higher cumulative confirmed cases rates.
- With the use of Jenks classification method, spatial patterns can be observed for all weeks. 
- In the earlier analysis using Equal Intervals classification method, it can be gathered that COVID-19 cases originated from the central region in February 2021. However, it can now be observed that initially in March 2020, some sub-districts located in the central region and along the western coast already have higher cumulative confirmed cases rates.
- It can also seen that the sub-districts in the central region are most hit by COVID-19, even across time. They are consistently represented by darker shades of blue in the choropleth maps, especially from August 2020 onwards. 
- Moreover, it can be understood that COVID-19 was spread to the southern region across time. This is because several sub-districts in the south are observed to have higher cumulative confirmed cases rates as the months pass.

### 5.4.2 Cumulative Death Rates By Month
```{r echo=TRUE, eval=TRUE}
tmap_arrange(jenks(jakartaDKI_rates, "MENINGGAL_Mar20"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Apr20"),
             jenks(jakartaDKI_rates, "MENINGGAL_May20"),
             jenks(jakartaDKI_rates, "MENINGGAL_Jun20"),
             jenks(jakartaDKI_rates, "MENINGGAL_Jul20"),
             jenks(jakartaDKI_rates, "MENINGGAL_Aug20"),
             jenks(jakartaDKI_rates, "MENINGGAL_Sep20"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Oct20"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Nov20"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Dec20"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Jan21"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Feb21"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Mar21"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Apr21"), 
             jenks(jakartaDKI_rates, "MENINGGAL_May21"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Jun21"), 
             jenks(jakartaDKI_rates, "MENINGGAL_Jul21"))
```

- A darker shade of blue signifies higher cumulative death rates.
- With the use of Jenks classification method, spatial patterns can be observed for all weeks. 
- In the earlier analysis using Equal Intervals classification method, it was observed that only 1 sub-district in the central region has higher cumulative death rates in February 2021. However, it can now be observed that initially in March 2020, some sub-districts located in the central region and along the north-western coast already have higher cumulative death rates.
- It can also seen that the sub-districts in the central region are most hit by COVID-19, even across time. They are consistently represented by darker shades of blue in the choropleth maps, especially from August 2020 onwards. 
- Moreover, it can be understood that COVID-19 was spread to the southern region across time. This is because several sub-districts in the south are observed to have higher cumulative death rates as the months pass.